### 2011-11-20 src/synopsis.tex
- corrected the synopsis template usage. - deleted chapters. made sections - basic org done. - the main changes and final! changes are pending.

\documentclass[MS,synopsis]{iitmdiss} 
\usepackage[usenames,dvipsnames]{color}

\usepackage{times} 
\usepackage{t1enc}
\usepackage{../lib/mylatexlib}

\title{\mythesistitle} 
\author{\myname} 
\date{\MakeUppercase
  {\mysubmissionmonth}} 
\department{\MakeUppercase{\mydept}}

\begin{document}

%\nocite{*}
\maketitle

% \tnote{TBD: DELETE IF WE RUN OUT OF SPACE}
% \begin{singlespace}
% \tableofcontents
% \thispagestyle{empty}
% \listoffigures
% \addcontentsline{toc}{chapter}{LIST OF FIGURES}
% \listoftables
% \addcontentsline{toc}{chapter}{LIST OF TABLES}
% \end{singlespace}

% The main text will follow from this point so set the page numbering
% to arabic from here on.
\pagenumbering{arabic}
\setcounter{page}{0}
%\newpage

\section{Introduction}
\label{sec:intro}

% \tnote{testing 1 2 3...}
% \tnote[bogus1]{testing 1 2 3...}
% \tnote[bogus2]{testing 1 2 3...}
% \tnote[bogus3]{testing 1 2 3...}
% \tnote[bogus4]{testing 1 2 3...}
% \tnote[bogus5]{testing 1 2 3...}
% \tnote[bogus6]{testing 1 2 3...}
% \tnote[bogus7]{testing 1 2 3...}
% \tnote[bogus8]{testing 1 2 3...}
% \tnote[TH]{testing 1 2 3...}

\subsection{Illustration of the problem}
\label{sec:problem}

% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
% \def \GGG {{\em ``{\bf G}irl with Curious Hair''}} 
% \def \BBB {{\em ``{\bf B}rief Interviews with Hideous Men''}}
% \def \TTT {{\em ``Everything and More: A Compact His{\bf T}ory of Infinity''}} 
% \def \xLLL {{\bf L}}
% \def \xGGG {{\bf G}} 
% \def \xBBB {{\bf B}}
% \def \xTTT {{\bf T}}

%{\bf L}aughing {\bf G}irl {\bf B}rief {\bf T}his

% \def \LLL {{\em ``{\bf L}aughing with Kafka''}} 
% \def \GGG {{\em ``Brief {\bf I}nterviews with Hideous Men''}} 
% \def \BBB {{\em ``Broom of the {\bf S}ystem''}}
% \def \TTT {{\em ``Everything and More: {\bf A} Compact History of Infinity''}}

% \def \xLLL {{\bf L}}
% \def \xGGG {{\bf I}} 
% \def \xBBB {{\bf S}}
% \def \xTTT {{\bf A}}

% \def \LLL {{\em ``{\bf B}rief Interviews with Hideous Men''}} 
% \def \GGG {{\em ``The String {\bf T}heory''}} 
% \def \BBB {{\em ``[{\bf W}]Rhetoric and the Math Melodrama''}}
% \def \TTT {{\em ``{\bf F}ate, Time, and Language: An Essay on Free Will''}}
\def \LLL {{\em ``$\mathbb{B}$rief Interviews with Hideous Men''}}
\def \GGG {{\em ``The String $\mathbb{T}$heory''}} 
\def \BBB {{\em ``[$\mathbb{W}$]Rhetoric and the Math Melodrama''}}
\def \TTT {{\em ``$\mathbb{F}$ate, Time, and Language: An Essay on Free Will''}}

% \def \xLLL {{\bf B}}
% \def \xGGG {{\bf T}} 
% \def \xBBB {{\bf W}}
% \def \xTTT {{\bf F}}
\def \xLLL {\mathbb{B}}
\def \xGGG {\mathbb{T}} 
\def \xBBB {\mathbb{W}}
\def \xTTT {\mathbb{F}}

%{{\em ``Everything and More: A Compact History of In{\bf F}inity''}} 
% {{\em ``Broom of {\bf T}he System''}} % {Consider the Lobster}
% {{\em ``God Bless You, Mr. {\bf F}ranzen''}}
% {{\em ``{\bf L}aughing with Kafka''}} 
% {{\em ``{\bf G}irl with Curious Hair''}} 
% {{\em ``Consider the {\bf L}obster|}}
% {{\em ``{\bf G}od Bless You, Mr. Franzen''}}
% {{\em ``{\bf T}ennis, Trigonometry,
%                                Tornadoes: A Midwestern  Boyhood''}}
% {{\em ``{\bf T}his is Water''}} %
% {\em {{\bf T}he Pale King}}

\def \Pa {{\em {\bf Pa}tricia}} 
\def \Pi {{\em {\bf Pi}gpen}} %{{\em {\bf Ma}rcie}}
\def \Sn {{\em {\bf Sn}oopy}}
\def \Wo {{\em {\bf Wo}odstock}}
\def \Vi {{\em {\bf Vi}olet}} 
\def \Li {{\em {\bf Li}nus}} 
\def \Ch {{\em {\bf Ch}arlie}}
\def \Sa {{\em {\bf Sa}lly}}
\def \Fr {{\em {\bf Fr}anklin}}  %{{\em {\bf Re}run}} 
\def \Sc {{\em {\bf Sc}hr{\"o}eder}} 
\def \Lu {{\em {\bf Lu}cy}}

\def \xPa {{\bf Pa}} 
\def \xPi {{\bf Pi}} 
\def \xSn {{\bf Sn}}
\def \xWo {{\bf Wo}}
\def \xVi {{\bf Vi}} 
\def \xLi {{\bf Li}} 
\def \xCh {{\bf Ch}}
\def \xSa {{\bf Sa}}
\def \xFr {{\bf Fr}}
\def \xSc {{\bf Sc}} 
\def \xLu {{\bf Lu}}

\def \residenceblock {{\em Infinite Loop}}

\def \WSI {{\em Wallace Studies Institute}}

\def \coneohone {{\em ``Influence of post modernism in Wallace's work''}}
\def \coneohtwo {{\em ``A study on fragmented prose method''}}

A group of students, \Pa, \Pi, \Sn, \Wo, \Vi, \Li, \Ch, \Sa, \Fr,
  \Sc\ and \Lu\ enroll at the
% {\bf Wo}odstock, {\bf Vi}olet, {\bf Li}nus, {\bf Ch}arlie, {\bf
%   Sa}lly, {\bf Re}run, {\bf Sc}hr{\"o}eder} and {\em {\bf Lu}cy}
% enroll at the
{\WSI} for a liberal arts programme.  As part
of their semester thesis, they pick a body of work to study and form
the namesake study groups, {\LLL}, {\GGG}, {\BBB} and {\TTT}. A student
will be in at least one study group and may be in more than one. For
instance, as will be seen later, {\Fr} studies both {\LLL} and {\TTT}
while \Wo\ studies only \BBB.

Let $U$ and $\cF$ represent the set of students and the set of study
groups respectively and the integers $n$ and $m$ denote the total
number students and study groups respectively. In relation to this
example, these are defined in Table~\ref{tab:wsigroups}. Also given
there is the study group allocation to students.

% \vspace{-3mm}
% \begin{align*}
% U &= \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
%     \xSa,\; \xFr,\; \xSc,\; \xLu\}\\
% \cF &= \{\xLLL, \xGGG, \xBBB, \xTTT\}\\
% \xLLL &= \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}\\
% \xGGG &= \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}\\
% \xBBB &= \{\xSn,\;  \xPi,\;  \xWo \}\\
% \xTTT &= \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}\\
% n &= |U| = 11\\
% m &= |\cF| = 4  
% \end{align*}

\begin{table}[htbp]
  \centering
  \begin{tabular}{rcl}
    $U $&$=$&$ \{\xPa,\; \xPi,\; \xSn,\; \xWo,\; \xVi,\; \xLi,\; \xCh,\;
    \xSa,\; \xFr,\; \xSc,\; \xLu\}$\\
    $\cF $&$=$&$ \{\xLLL, \xGGG, \xBBB, \xTTT\}$\\
    $\xLLL $&$=$&$ \{\xCh,\;  \xSa,\;  \xFr,\;  \xSc,\;  \xLu \}$\\
    $\xGGG $&$=$&$ \{\xPa,\;  \xPi,\;  \xVi,\;  \xCh \}$\\
    $\xBBB $&$=$&$ \{\xSn,\;  \xPi,\;  \xWo \}$\\
    $\xTTT $&$=$&$ \{\xVi,\;  \xLi,\;  \xCh,\;  \xFr \}$\\
    $n $&$=$&$ |U| = 11$\\
    $m $&$=$&$ |\cF| = 4$      
  \end{tabular}
  \caption{Students and study groups in \WSI}
  \label{tab:wsigroups}
\end{table}

% The administration faces a small challenge with their
% accommodation. 
The campus has a residential area {\residenceblock} that has $n$
single occupancy apartments reserved for the study groups'
accommodation.  All these apartments are located such that the streets
connecting them do {\em not} form loops. Fig~\ref{fig:streetmap} shows
the street map for {\residenceblock}. It may be noted that as a graph, it
classifies as a tree.

A natural question would be to find how the students should be
allocated apartments such that each study group has the {\em least
  distance to travel} for a discussion? More specifically, we are
interested in enforcing additional conditions, namely, that all the
students in a study group must be next to each other; in other words,
for one student to reach another fellow study group member's apartment
(for all study groups the student is part of), she must not have to
pass the apartment of any student who is not in that study group. To
further elucidate, the apartments of students of any study group must
be arranged in an exclusive unfragmented path on the street
map. Exclusivity here means that the path must not have apartments
from other study groups (unless that apartment is also part of {\em
  this} study group).

An intuitive approach to this problem would be to first find the paths
that each study group decides to inhabit and then refine the
allocation to individual students. A feasible allocation of exclusive
routes to study groups is illustrated in Fig~\ref{fig:streetmappath}
and the students' allocation of apartments that obeys this route
allocation is shown in Fig~\ref{fig:streetmappathpeople}. How this is
algorithmically computed is the focus of this thesis.

\begin{figure}[htbp]
 \centering
  \includegraphics[scale=0.3]{../img/1_infinite_loop.png}
  \caption{{\residenceblock} street map}
  \label{fig:streetmap}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{../img/2_infinite_loop_BTWF.png}
  \caption{{\residenceblock} street map with study group routes
    allocated. Routes are color coded as follows: red for
    \textcolor{red}{$\xLLL$} group, blue for \textcolor{blue}{$\xGGG$}
    group, orange for \textcolor{YellowOrange}{$\xBBB$} group, green for
    \textcolor{green}{$\xTTT$} group}
  \label{fig:streetmappath}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../img/3_infinite_loop.png}
  \caption{Individual allocation of apartments to students in
    {\residenceblock} that meets the requirements stated before.}
  \label{fig:streetmappathpeople}
\end{figure}

As a special case, suppose all the apartments are on the same street
or if they are all lined up on a single path, the street map becomes a
tree that is just a path. Then the problem becomes what is called an
{\em interval assignment problem}. The idea of interval assignment may
not be obvious here; hence to see this, consider a different problem
in {\WSI} where the classes for these study groups courses need to be
scheduled during a day (or a week or any time period). Each study
group has a bunch of courses associated with it some of which may be
shared by two or more study groups. It is mandatory that a student who
is a member of a study group takes all the courses associated with
that group. There are slots during the day for classes to be held and
the problem is to allocate class slots to courses such that all the
classes of a study group are consecutive. It is debatable if this will
not hamper the attention span and memory retention rate of the
students but that is, regrettably, out of the scope of this
thesis. The parallels between this class allocation problem and the
accommodation problem can be seen as follows. The set $U$ here, are
the courses offered (say Course 101 {\coneohone}, Course 102
{\coneohtwo} and so on). In this variation of the problem, the
collection $\cF$ is the set of study groups but the study groups are
filled by course IDs (in place of students in the earlier
example). For instance, Course 101 is mandatory for all study groups
$\xLLL$, $\xGGG$, $\xBBB$, $\xTTT$ and Course 102 is mandatory for
only the $\xLLL$ group) and so on. The sequence of class slots for the
day (or week or any time period) is analogous to the street map in the
accommodation problem. It is quite obvious now why this version of the
problem (where the ``target graph'' is a path and not any
tree\tnote[TH]{Allowing any tree in this example could be seen as a
  scenario where there are parallel classes. A node falling in the
  path between two other nodes would mean that the corresponding is
  scheduled between the other two.}) is called an interval assignment
problem.

The interval assignment problem to a set system is equivalent to the
consecutive ones problem (COP) in binary matrices\cite{wlh02,
  nsnrs09}.  The COP problem is to rearrange rows (columns) of a
binary matrix in such a way that every column (row) has its $1$s occur
consecutively. If this is possible the matrix is said to have the COP.
Consecutive ones property (COP) of binary matrices is a well
researched combinatorial problem and has several positive results on
tests for it and computing the COP permutation (i.e. the class slot
allocation to course IDs in the above illustration) which will be
surveyed later in this document. Hence we are interested in extensions
of COP, more specifically, the extension of interval assignment
problem to tree path assignment problem (which is illustrated by the
study group accommodation problem).

\tnote{add a note on the organization of the doc}
% The necessary preliminaries with definitions etc. are presented in
% Section~\ref{sec:prelims}. 
% Section~\ref{sec:feasible} documents the
% characterization of a feasible path labeling and finally,
% Section~\ref{sec:ksubdivstar} describes a polynomial time algorithm to
% find the tree path labeling of a given set system from a given
% $k$-subdivided tree.

Section~\ref{sec:motivation} gives a brief survey of COP and relatives.
Section~\ref{sec:results} presents a summary of our results on the
tree path labeling problem.

\section{Motivation and background}
\label{sec:motivation}

As described earlier, a matrix with COP is one whose rows (columns)
can be rearranged so that the 1s in every column (row) are in
consecutive rows (columns).

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.4]{../img/0_cop_nocop.png}

  \caption{{\em Example for the COP:} The matrix on the left has the
    COP because by permuting its columns (labeled with $c_1$-$c_4$)
    one can obtain the matrix shown in the middle where the 1s in each
    row appear consecutively. The matrix on the right, in contrast,
    does not have the COP \cite{at72,d08phd}.}
  \label{fig:cop-matrix}
\end{figure}

COP in binary matrices has several practical applications in diverse
fields including scheduling \cite{hl06}, information retrieval
\cite{k77} and computational biology \cite{abh98}.  Further, it is a
tool in graph theory \cite{mcg04} for interval graph recognition,
characterization of Hamiltonian graphs, planarity testing \cite{bl76}
and in integer linear programming \cite{ht02,hl06}.

The obvious first questions after being introduced to the consecutive
ones property of binary matrices are if COP can be detected
efficiently in a binary matrix and if so, can the COP permutation of
the matrix also be computed efficiently?  Recognition of COP in a
binary matrix is polynomial time solvable and the first such algorithm
was given by \cite{fg65}.  Since then several algorithms have been
invented; PQ trees \cite{bl76} gave the first linear time
algorithm. Variations of PQ trees \cite{mm96,wlh01,wlh02,mcc04} and
ICPIA \cite{nsnrs09} further improve either the time complexity and/or
the complexity of implementation.

\cite{bl76} describes a linear time construction of a data structure
called PQ trees. This paper draws on the close relationship of COP
matrices to interval graphs. A PQ tree of a matrix is one that stores
all row (column) permutations of the matrix that give the COP orders
(there could be multiple) of the matrix. This is constructed using an
elaborate linear time procedure. Since this result, several
improvements have been made to the procedural complexity of the
algorithm itself and also derivative data structures that are easier
to construct. PQR trees is one such data structure \cite{mm96},
\cite{mpt98}\tnote[TH]{also, mpt98 $==$ elaborate version of mm96}.
\cite{tm05} describes an improved version of PQR
trees. \tnote{improv in terms of what?}\cite{wlh02} describes the
simpler algorithm for COT. Hsu also invented PC trees
\cite{wlh01}\tnote[TH]{This result first appeared inproc ISAAC92}
which is claimed to be much easier to implement. \cite{nsnrs09}
describes a characterization of consecutive ones property solely based
on the cardinality properties of the set representations of the
columns (rows); every column (row) is equivalent to a set that has the
row (column) indices of the rows (columns) that have one entries in
this column (row). This is interesting and relevant, especially to
this thesis because it simplifies COT to a great
degree. \tnote{get this
  figured out! - because it reduces the search space for the solution
  from set of all permutations of rows to set of all}

\cite{mcc04} describes a different approach to COT. While all previous
COT algorithms gave the COP order if the matrix has the property and
if not it exited stating negative, this algorithm gives an evidence by
way of a certificate of
matrix being devoid of COP thereby enabling a user to check the
algorithm's result. This is significant from an implementation
perspective because automated program verification is hard and manual
verification is more viable. Hence having a certificate
reinforces its credibility. Note that when the matrix has COP the COP
order is the certificate.  It may also be noted that the internal
machinery of this algorithm \annote {is related to the weighted
  betweenness problem addressed } {in what way??} in \cite{co98}.
\tnote{expand on the COP order graph creation and it having to be
  bipartite for M to have COP. and thus an odd cycle being an evidence
  of no COP.}
\tnote[TH]{ where should this go?: cite|jlm97 (application of PQ trees
  in graphics).\\ add helly's theorem citation
  19XXdgk-Hellystheorem-Danzer-Gruenbaum-Klee}

The literature also has results on parallel COT algorithms --
\cite{as95} \cite{bs03} \cite{ly91}\tnote{add english to fit context}

\subsection{Optimization problems in COP}
So far we have been concerned about
matrices that have the consecutive ones property. However in real life
applications, it is rare that data sets represented by binary matrices
have COP. Nevertheless, the COP is not arbirary and there are several
interesting problems when a matrix does not have COP but is ``close''
to having COP or can be altered to have COP.

-- sect 4.1 in \cite{d08phd} has many results surveyed. hardness results,
approx. results. results are usually for a class of matrices $(a,b)$
where number columns and rows are restriced to $a$ and $b$
\tnote{check if it is the other way around}.\\
-- problem of flipping at most $k$ entries of $M$ to make it attain
COP. this is NP complete \cite{b75-phd}

The first step in this direction is to analyze matrices without
COP. \cite{at72} showed that a matrix that does not have COP have
certain forbidden substructures that prevent it from having
COP. Tucker classified these substructures into five classes of
submatrices. This result is presented in the context of AT-free
bipartite graphs \tnote[TH]{ check this up. give details.} which is
equivalent to matrices with COP. A binary matrix is converted into a
bipartite graph as follows. The set of rows and the set of columns
form the two partitions of the graph. Each row node is adjacent to
those nodes that represent the columns that have ones in the
corresponding row. \cite{at72} proves that this bipartite graph has no
asteroidal triple if and only if the matrix has COP and goes on to
identify the forbidden substructures for these bipartite graphs. The
matrices corresponding to these substructures are the forbidden
submatrices. Once a matrix has been detected to not have COP (using
any of the COT algorithms mentioned earlier), it is naturally of
interest to find out the smallest (in terms of number of rows, columns
and number of entries that are ones) forbidden
substructure. \cite{d08phd} discusses an exact algorithm with time
complexity \annote{parameterized by the number of ones in the
  matrix}{check this.}.

{\sc Min COS-R (Min COS-C), Max COS-R (Max COS-C)} are similarly
inspired problems. In {\sc Min COS-R (Min COS-C)} the question is to
find the minimum number of rows (columns) that must be deleted to
result in a matrix with COP.  In the dual problem {\sc Max COS-R (Max
  COS-C)} the search is for the maximum number of rows (columns) that
induces a submatrix with COP. Given a no-COP matrix $M$,
\cite{b75-phd} shows that finding a submatrix $M'$ with all columns
but a maxmimum cardinality subset of rows such that $M'$ has COP is NP
complete\tnote{check if b75 deals with COP col or COP row. also
  is it any submatrix with k less than r rows or submatrix must have
  all columns?}. \cite{hg02} corrects an error of the abridged proof
of this reduction as given in \cite{gj79}.  \cite{d08phd} discusses
all these problems in detail giving an extensive survey of the
previously existing results which are almost exhaustively all
approximation results and hardness results. Taking this further,
\cite{d08phd} presents new results in the area of parameterized
algorithms for this problem\tnote[TH]{elaborate - what are
  the results?}.

Another problem is to find the minumum number of entries in the matrix
that can be flipped to result in a matrix with COP.  \cite{v85}
discusses approximation of {\sc COP Augmentation} which is the problem
of changing of the minimum number of zero entries to ones so that the
resulting matrix has COP. As mentioned earlier, this problem had been
known to be NP complete due to \cite{b75-phd}. \cite{v85} also proves,
using a reduction to the longest path problem, \tnote{or is it a
  survey of another result?  check.} that finding a Tucker's forbidden
submatrix of at least $k$ rows is NP complete. \tnote{how is this
  different from booth's 75 result??}  \tnote[TH]{where should this
  go? cite|tz04 (approx submatrix with COP sparse matrices)}

\cite{jkckv04} discusses the use of matrices with almost-COP in the
storage of very large databases.  The problem is that of $k$-run
reordering of binary matrices which they prove to be NP hard using a
reduction from the hamiltonian path problem.  A connection of COP
problem to the travelling salesman problem is also
introduced. \tnote[TH]{what does this mean? -- COP can be used as a
  tool to reorder $0.5T \le runs(M) \le T$}

The optimization version of the $k$-run problem, i.e. minimization of
number of blocks of ones is proven to be NP complete by
\cite{k77}. \tnote[TH]{what is the reduction?}

\tnote[TH]{other problems similar to COP -- cite:ckl96 (ILP, circ ones, one drop) -- cite:th98 (generalization of COP - minimax, biotonic column) Tucker}

{\bf Motivation: } A natural generalization of the interval assignment problem is
feasible tree path labeling problem of a set system. The problem is
defined as follows -- given a set system $\cF$ from a universe $U$ and
a tree $T$, does there exist a bijection from $U$ to the vertices of
$T$ such that each set in the system maps to a path in $T$.  We refer
to this as the {\em tree path labeling problem} for an input set
system, target tree pair -- $(\cF,T)$. As a special case if the tree
$T$ is a path, the problem becomes the interval assignment problem.
We focus on the question of generalizing the notion of an ICPIA
\cite{nsnrs09} to characterize feasible path assignments.  We show
that for a given set system $\cF$, a tree $T$, and an assignment of
paths from $T$ to the sets, there is a feasible bijection between $U$
and $V(T)$ if and only if all intersection cardinalities among any
three sets (not necessarily distinct) is same as the intersection
cardinality of the paths assigned to them and the input passes a
filtering algorithm (described in this paper) successfully.  This
characterization gives a natural data structure that stores all the
relevant feasible bijections between $U$ and $V(T)$. This reduces the
search space for the solution considerably from the universe of all
possible bijections between $U$ and $V(T)$ to only those bijections
that maintain the characterization\tnote{REWORD}.  Further, the
filtering algorithm is also an efficient algorithm to test if a tree
path labeling to the set system is feasible.  This generalizes the
result in \cite{nsnrs09}.

Our results also have close connection to recognition of path graphs
and connection to path graph isomorphism.  Graphs which can be
represented as the intersection graph of paths in a tree are called
{\em path graphs}\cite{mcg04}. Thus given a hypergraph $\cF$, it can
be viewed as paths in a tree, if and only if the intersection graph of
$\cF$ is a {\em path graph}. Path graphs are a subclass of chordal
graphs since chordal graphs are combinatorially characterized as the
intersection graphs of subtrees of a tree.  Path graphs are well
studied in the literature \cite{plr70,gav78,bp93,mcg04}.  Checking if
a graph is a path graph can be done in polynomial time by the results
of \cite{gav78,aas93}.  However, this is only a necessary condition
for our question.  \remove{Indeed, our work provides a sufficient
  condition for a related problem: that is whether a path assignment
  to be feasible?}  On the other hand, path graph isomorphism is known
be isomorphism-complete; see for example \cite{kklv10}. Therefore,
\annote{it is unlikely that we can solve the problem of finding
  feasible path labeling $\cl$ for a given $\cF$ and }{WHY?} tree $T$.
\annote{It is definitely interesting to classify the kinds of trees
  and hypergraphs for which feasible path labelings can be
}{ELABORATE.} found efficiently.  \annote{These results would form
  a natural generalization of COP testing and interval graph
  isomorphism, culminating in Graph Isomorphism }{REWORD} itself.

How the problem of COT is related to the constraint statisfaction
problem of interval assignment to sets in a system can be seen as
follows. Every column of the binary matrix can be converted into a set
of integers which are the indices of rows with $1$s in that column. It
is apparent that if the matrix has COP, then constructing such sets
after applying the COP permutation to the matrix will result in all
sets with consecutive integers. In other words, after application of
COP reordering, the sets are intervals. Indeed the problem now becomes
finding interval assignments to a given set system such that there
exists a permutation of the universe (set of row indices) which
converts each set to its assigned interval. The result in
\cite{nsnrs09} characterizes interval assignments to the sets which
can be obtained from a single permutation of the rows.  They show that
for each set, the cardinality of the interval assigned to it must be
same as the cardinality of the set, and the intersection cardinality
of any two sets must be same as the intersection cardinality of the
corresponding intervals.  While this is obviously a necessary
condition, this result shows this is also sufficient.  \cite{nsnrs09}
calls such an interval assignment an Intersection Cardinality
Preserving Interval Assignment (ICPIA).  This paper generalizes the
idea from \cite{wlh02} of decomposing a given binary matrix into prime
matrices for COT and describes an algorithm to test if an ICPIA exists
for a given set system.

\section{Summary of results}
\label{sec:results}

Given a path labeling $\cl$ to $\cF$ from a tree $T$, we give a
necessary and sufficient condition for it to be a feasible path
labeling.  This necessary and sufficient condition can be tested in
polynomial time.  The most interesting consequence is that in our
constructive procedure, it is sufficient to iteratively check if
three-way intersection cardinalities are preserved.  In other words,
in each iteration, it is sufficient to check if the intersection of
any three sets is of the same cardinality as the intersection of the
corresponding paths.  This generalizes the well studied question of
the case when the given tree $T$ is a path \cite{wlh02,nsnrs09}.

%\item 
In the Section 4, \annote{we initiate an exploration of}{WTF?}
finding feasible path labeling of set systems in a special kind of
tree which we call the $k$-subdivided star.  \annote{This question
  is an attempt to generalize}{REWORD.} the problem of testing if a
matrix has the consecutive ones property.  However, we restrict the
hypergraph $\cF$ to be such that all hyperedges have at most $k+2$
elements.  \annote{In spite of this restricted case we consider, we
  believe that our results are of significant interest in
  understanding the nature of Graph Isomorphism}{REWORD.} which is
polynomial time solvable in interval graphs and is hard on path
graphs.
%\end{enumerate}

% In the later part of this paper, we decompose our search for a
% bijection between $U$ and $V(T)$ into subproblems.  Each subproblem
% is on a set subsystem in which for each set, there is another set in
% the set subsystem with which the intersection is {\em strict}, i.e.,
% there is a non-empty intersection, but neither is contained in the
% other.  This is in the spirit of results in \cite{wlh02,nsnrs09}
% where to test for the COP in a given matrix, the COP problem is
% solved on an equivalent set of prime matrices.

%\subsection{RESULTS}
% \section{Preliminaries} 
% \label{sec:prelims}

% \subsection{Hypergraph} 
% The set $\F \subseteq (2^{U} \setminus \emptyset)$ is a {\em set
%   system} of a universe $U$ with $|U| = n$.  The {\em support} of a
% set system $\F$ denoted by $supp(\cF)$ is the union of all the sets in
% $\F$; $supp(\F) = \bigcup_{S \in \F}S$. For the purposes of this
% paper, a set system is required to ``cover'' the universe; $ supp(\cF)
% = U$.
% % A set system $\cF$ can be alternatively represented by a {\em
% %   hypergraph} $\H_\cF$ whose vertex set is $supp(\cF)$ and
% % hyperedges are the sets in $\cF$. This is a known representation for
% % interval systems in literature \cite{bls99,kklv10}.  We extend this
% % definition here to path systems.
% A set system $\cF$ can also be visualized as a {\em
%   hypergraph}\, vertex set is $supp(\cF)$ and hyperedges are the sets
% in $\cF$.  The {\em intersection graph}\, $\bI(\cF)$ of a hypergraph
% $\cF$ is a graph such that its vertex set has a bijection to $\cF$ and
% there exists an edge between two vertices \iff their corresponding
% hyperedges have a non-empty intersection \cite{mcg04}.  Two
% hypergraphs $\cF'$, $\cF''$ are said to be {\em isomorphic} to each
% other, denoted by $\cF' \cong \cF''$, \iff there exists a bijection
% $\phi: supp(\cF') \rightarrow supp(\cF'')$ such that for all sets $A
% \subseteq supp(\cF')$, $A$ is a hyperedge in $\cF'$ \iff  $B$ is a
% hyperedge in $\cF''$ where $B = \{\phi(x) \mid x \in A\}$
% \cite{kklv10}, written as $B=\phi(A)$.
% % This is called {\em hypergraph isomorphism}.

% %\noindent
% \subsection{Path Hypergraph from a Tree} 
% We consider trees $T$ such that $|V(T)|=|U|=n$.  A {\em path system}\,
% $\cP$ is a set system of paths from $T$; $\cP \subseteq \{P \mid P
% \subseteq V, \text{ } T[P] \text{ is a path} \}$.  This generalizes
% the fact, from the literature \cite{bls99,kklv10}, that intervals can
% be viewed as sub-paths of a path.
% % Due to the equivalence of set system and hypergraph in the scope of
% % this paper, we drop the subscript $_H$ in the notation and refer to
% % both the structures by $\cF$.
% % \vspace{\prelimspace}%\tnote{also extend $\phi$ to hyperedges -- see if
% % required } \vspace{\prelimspace} \vspace{\prelimspace}
% % \noindent
% If the intersection graphs of $\cF$ and $\cP$( a path system) are
% isomorphic, $\bI(\cF) \cong \bI(\cP)$, then the associated bijection
% $\cl: \cF \rightarrow \cP$ due to this isomorphism is called a {\em
%   path labeling} of the hypergraph $\cF$.  Note that there are two
% kinds of isomorphisms here.  We are concerned about the isomrophisms
% intersection graphs on $\cF$ and $\cP$, and also the isomorphism
% between the hypergraph $\cF$ and $\cP$.
% % To illustrate further, let $\cg: V(\cF) \rightarrow V(\cP)$ be the
% % above mentioned isomorphism where $V(\cF)$ and $V(\cP)$ are the
% % vertex sets that represent the hyperedges for each hypergraph
% % respectively, $V(\cF) = \{ v_S \mid S \in \cF\}$ and $V(\cP) = \{
% % v_P \mid P \in \cP\}$. Then the path labeling $\cl$ is defined as
% % follows: $\cl(S_1) = P_1$ \iff $\cg (v_{S_1}) = v_{P_1}$.  Just to
% % emphasize, for a path labeling $\cl$ of $\cF$ with $\cP$ as the path
% % system, $\cF^\cl$ is same as $\cP$.  The path system $\cP$ may be
% % alternatively denoted in terms of $\cF$ and $\cl$ as $\cF^\cl$. In
% % most scenarios in this paper, what is given are the pair $(\cF,
% % \cl)$ and the target tree $T$; hence this notation will be used more
% % often.
% If $\cF \cong \cP$ where $\cP$ is a path system, then $\cF$ is called
% a {\em path hypergraph} and $\cP$ is called {\em path representation}
% of $\cF$. If this isomorphism is $\phi: supp(\cF) \rightarrow V(T)$,
% then it is clear that there is a path labeling $\cl_\phi: \cF
% \rightarrow \cP$ to the set system; $\cl_\phi(S) = \set{y \mid y =
%   \phi(x), x \in S}$ for all $S \in \cF$. In other words, if $\cF
% \cong \cP$, we get a path labeling.  Recall that $supp(\cP) = V(T)$.
% % \vspace{\prelimspace} A graph $G$ is a {\em path graph} if it is
% % isomorphic to the intersection graph $\bI(\cP)$ of a path system
% % $\cP$.  This isomorphism gives a bijection $\cl': V(G) \rightarrow
% % \cP$. Moreover, for the purposes of this paper, we require that in a
% % path labeling, $supp(\cP) = V(T)$.  If graph $G$ is also isomorphic
% % to $\bI(\cF)$ for some hypergraph $\cF$, then clearly there is a
% % bijection $\cl: \cF \rightarrow \cP$ such that $\cl(S) = \cl'(v_S)$
% % where $v_S$ is the vertex corresponding to set $S$ in $\bI(\cF)$ for
% % any $S \in \cF$. This bijection $\cl$ is called the {\em path
% %   labeling} of the hypergraph $\cF$ and the path system $\cP$ may be
% % alternatively denoted as $\cF^\cl$.  \vspace{\prelimspace}
% In this work, we are given as input $\cF$ and a tree $T$, and the
% question is whether there is a path labeling $\cl$ to a set of paths
% in $T$.  We refer to such a solution path system by $\cF^\cl$.  A path
% labeling $(\cF, \cl)$ is defined to be {\em feasible} if
% % $\cF \cong \cF^\cl$ and this
% there is a hypergraph isomorphism $\phi: supp(\cF) \rightarrow
% supp(\cF^\cl)=V(T)$ induces a path labeling $\cl_\phi: \cF \rightarrow
% \cF^\cl$ such that $\cl_\phi = \cl$.
% % \vspace{\prelimspace}

% \subsection{Overlap Graphs and Marginal Hyperedges} 
% An {\em
%   overlap graph}\, $\bO(\cF)$ of a hypergraph $\cF$ is a graph such
% that its vertex set has a bijection to $\cF$ and there exists an edge
% between two of its vertices \iff their corresponding hyperedges
% overlap. Two hyperedges $S$ and $S'$ are said to {\em overlap},
% denoted by $S \overlap S'$, if they have a non-empty intersection and
% neither is contained in the other; $S \overlap S' \text{ iff } S \cap
% S' \ne \emptyset, S \nsubseteq S', S' \nsubseteq S$. Thus $\bO(\cF)$
% is a spanning subgraph of $\bI(\cF)$ and not necessarily
% connected. Each connected component of $\bO(\cF)$ is called an {\em
%   overlap component}.
% % If there are $d$ overlap components in $\bO(\cF)$, the set
% % subsystems are denoted by $\cO_1, \cO_2, \ldots \cO_d$. Clearly
% % $\cO_i \subseteq \F, i \in [d]$. For any $i, j \in [d]$, it can be
% % verified that one of the following is true.
% % \begin{enumerate}[a) ]
% % \item $supp(\cO_i)$ and $supp(\cO_j)$ are disjoint
% % \item $supp(\cO_i)$ is a subset of a set in $\cO_j$
% % \item $supp(\cO_j)$ is a subset of a set in $\cO_i$
% % \end{enumerate}
% % \vspace{\prelimspace} \xnoindent
% A hyperedge $S \in \cF$ is called {\em marginal} if for all $S'
% \overlap S$, the overlaps $S \cap S'$ form a single inclusion chain
% \cite{kklv10}. Additionally, if $S$ is such that it is contained in no
% other marginal hyperedge in $\cF$, then it is called {\em
%   super-marginal}.
% % i.e., it is inclusion maximal then it is called {\em
% %   super-marginal}.

% \subsection{$k$-subdivided Star -- A Special Tree} 
% A {\em star} graph
% is a complete bipartite graph $K_{1,p}$ which is clearly a tree and
% $p$ is the number of leaves. The vertex with maximum degree is called
% the {\em center} of the star and the edges are called {\em rays} of
% the star. A {\em $k$-subdivided star} is a star with all its rays
% subdivided exactly $k$ times. The definition of a {\em ray of a
%   $k$-subdivided star} is extended to the path from the center to a
% leaf. It is clear that all rays are of length $k+2$. See
% fig.~\ref{fig:kstar} for examples.

% \begin{figure}[htbp] 
%   \centering
%   (i) \includegraphics{cop-tree_thesis.1} \ \ \ \ \ \  
%   (ii)  \includegraphics{cop-tree_thesis.2}
%   \caption{(i) $8$-subdivided star with 7 rays, (ii) 3-subdivided
%     star with 3 rays}
%   \label{fig:kstar}
% \end{figure}

\subsection{Characterization of Feasible Tree Path  Labelings} \tnote[TH]{UNCOMMENT all proofs}
\label{sec:feasible}

In this section we give an algorithmic characterization of a
feasibility of tree path labeling.
% Consider a path labeling $\cl: \cF \rightarrow \cP$ for set system $\cF$
% and path system $\cP$ on the given tree $T$. We
% call $\cl$ an {\em Intersection Cardinality Preserving Path Labeling
%   (ICPPL)} if it has the following properties.
Consider a path labeling $(\cF, \cl)$ on the
given tree $T$. We call $(\cF, \cl)$ an {\em Intersection Cardinality
  Preserving Path Labeling (ICPPL)} if it has the following
properties.

% \begin{enumerate}[{\icpplpr\ }i. ]
% \item \label{pr:i} $|S| = |\cl(S)|$, \hfill for all $S \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:ii}$|S_1 \cap S_2| = |\cl(S_1) \cap \cl(S_2)|$, \hfill
%   for all distinct $S_1, S_2 \in \cF$
%   % \vspace{\topshrink}
% \item \label{pr:iii}$|S_1 \cap S_2 \cap S_3| = |\cl(S_1) \cap \cl(S_2)
%   \cap \cl(S_3)|$, \hfill for all distinct $S_1, S_2, S_3 \in \cF$
% \end{enumerate}

\begin{table}[h]
  \centering
  \begin{tabular}[l]{rrcll}
    Pr i. & \label{xpr:i} $|S| $&$=$&$ |\cl(S)|$ & $\forall$ $S \in \cF$\\
    Pr ii. & \label{xpr:ii}$|S_1 \cap S_2| $&$=$&$ |\cl(S_1) \cap
    \cl(S_2)|$ &  $\forall$ distinct $S_1, S_2 \in \cF$\\
    Pr iii. & \label{xpr:iii}$|S_1 \cap S_2 \cap S_3| $&$=$&$
    |\cl(S_1) \cap \cl(S_2) \cap \cl(S_3)|$ & $\forall$ distinct $S_1,
    S_2, S_3 \in \cF$ 
  \end{tabular}
  \caption{ICPPL properties}
  \label{tab:icpplpr}
\end{table}

\noindent
The following lemma is useful in subsequent arguments. 
\begin{lemma}
  \label{lem:setminuscard}
  If $\cl$ is an ICPPL, and $S_1, S_2, S_3 \in \cF$, then $|S_1 \cap
  (S_2 \setminus S_3)| = |\cl(S_1) \cap (\cl(S_2) \setminus
  \cl(S_3))|$.
\end{lemma}

% \begin{proof}%[Proof of Lemma~\ref{lem:setminuscard}]
%   Let $P_i = \cl(S_i)$, for all $i \in \set{1,2,3}$. We have $|S_1
%   \cap (S_2 \setminus S_3)| = |(S_1 \cap S_2) \setminus S_3| = |S_1
%   \cap S_2| - |S_1 \cap S_2 \cap S_3|$. Due to properties
%   (\ref{pr:ii}) and (\ref{pr:iii}) of ICPPL, $|S_1 \cap S_2| - |S_1
%   \cap S_2 \cap S_3| = |P_1 \cap P_2| - |P_1 \cap P_2 \cap P_3| =
%   |(P_1 \cap P_2) \setminus P_3| = |P_1 \cap (P_2 \setminus
%   P_3)|$. Thus lemma is proved. \qed
% \end{proof}

In the remaining part of this section we show that $(\cF, \cl)$ is
feasible if and only if it is an ICPPL and
Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
function. Algorithm~\ref{al:icppl-find-isomorph} recursively does two
levels of filtering of $(\cF, \cl)$ to make it simpler while retaining
the set of isomorphisms, if any, between $\cF$ and $\cF^\cl$.
% One direction of this claim isclear: that if a path
% labeling is feasible, then all intersection cardinalities are
% preserved, i.e. the path labeling is an ICPPL. Algorithm~\ref{perms}
% \annote{has no premature exit condition hence any input will go
%   through it}{Prove that the filtered sets has ICPPL \iff input PL
%   has ICPPL?}. Algorithm~\ref{leafasgn} has an exit condition at
% line~\ref{xempty}. It can be easily verified that $X$ cannot be
% empty if $\cl$ is a feasible path labeling. The reason is that a
% feasible path labeling has an associated bijection between
% $supp(\cF)$ and $V(T)$ \remove{i.e. $supp(\cF^{\cl})$} such that
% the sets map to paths, ``preserving'' the path labeling.  The rest
% of the section is devoted to constructively proving that it is
% sufficient for a path labeling to be an ICPPL and pass the two
% filtering algorithms.  To describe in brief, the constructive
% approaches refine an ICPPL iteratively, such that at the end of each
% iteration we have a ``filtered'' path labeling, and finally we have
% a path labeling that defines a family of bijections from $supp(\cF)$
% to $V(T)$\remove{ i.e. $supp(\cF^{\cl})$}.

First, we present Algorithm~\ref{perms} called {\tt filter\_1}
% , and
% prove its correctness.  
This algorithm refines the path labeling by
processing pairs of paths in $\cF^\cl$ that share a leaf until no two
paths in the new path labeling share any leaf.
\tnote{write a note summarizing filter 1}
% \begin{algorithm}[h]
%   \caption{Refine ICPPL {\tt filter\_1($\cF, \cl, T$)}}
%   \label{perms}
%   \begin{algorithmic}[\lndisplay]
%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$\\
%     \STATE $j \assign 1$\\
%     \WHILE {there is $S_1, S_2 \in \cF_{j-1}$ such that
%       $\cl_{j-1}(S_1)$ and $\cl_{j-1}(S_2)$ have a common leaf in
%       $T$}\label{shareleaf} \STATE $\cF_j \assign (\cF_{j-1} \setminus
%     \{S_1, S_2\})
%     \cup \{S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1 \}$ \label{setbreak} 
%     \COMMENT {Remove $S_1$, $S_2$ and add the ``filtered'' sets}
%     \STATE {\bf for} every $S \in \cF_{j-1}$ s.t. $S \ne S_1$ and $S \ne
%     S_2$ {\bf do} $\cl_j(S) \assign \cl_{j-1}(S)$ {\bf end for}\\

%     \STATE $\cl_j(S_1 \cap S_2) \assign \cl_{j-1}(S_1) \cap
%     \cl_{j-1}(S_2)$
%     \COMMENT {Carry forward the path labeling for all existing sets other than
%       $S_1$, $S_2$}
%     \STATE $\cl_j(S_1 \setminus S_2) \assign \cl_{j-1}(S_1) \setminus
%     \cl_{j-1}(S_2)$ 
%     \COMMENT {Define path labeling for new sets}
%     \STATE $\cl_j(S_2 \setminus S_1) \assign \cl_{j-1}(S_2) \setminus
%     \cl_{j-1}(S_1)$

%     \IF{$(\cF_j, \cl_j)$ does not satisfy (\icpplpr~\ref{pr:iii}) of ICPPL}
%     \label{ln:3waycheck}
%     \STATE {\bf exit} \label{ln:exit1} \\
%     \ENDIF

%     \STATE $j \assign j+1$\\
%     \ENDWHILE
%     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
%     \RETURN $(\cF', \cl')$
%   \end{algorithmic}
% \end{algorithm}

\begin{lemma} 
 \label{lem:feasible} 
 In Algorithm~\ref{perms}, if input $(\cF, \cl)$ is a feasible path
 assignment then at the end of $j$th iteration of the {\bf while}
 loop, $j \ge 0$, $(\cF_j, \cl_j)$ is a feasible path assignment.
\end{lemma}

% \begin{proof}%[Proof of Lemma~\ref{lem:feasible}]
%   We will prove this by mathematical induction on the number of
%   iterations. The base case $(\cF_0, \cl_0)$ is feasible since it is
%   the input itself which is given to be feasible. Assume the lemma is
%   true till $j-1$th iteration. i.e. every hypergraph isomorphism
%   $\phi: supp\left(\cF_{j-1}\right) \rightarrow V\left(T \right)$ that
%   defines $(\cF, \cl)$'s feasibility, is such that the induced path
%   labeling on $\cF_{j-1}$, $\cl_{\phi[{\cF_{j-1}}]}$ is equal to
%   $\cl_{j-1}$. We will prove that $\phi$ is also the bijection that
%   makes $(\cF_j, \cl_j)$ feasible. Note that $supp(\cF_{j-1}) =
%   supp(\cF_{j})$ since the new sets in $\cF_j$ are created from basic
%   set operations to the sets in $\cF_{j-1}$. For the same reason and
%   $\phi$ being a bijection, it is clear that when applying the $\phi$
%   induced path labeling on $\cF_j$, $ \cl_{\phi[{\cF_{j}}]}(S_1
%   \setminus S_2) = \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
%   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Now observe that $ \cl_j(S_1
%   \setminus S_2) = \cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2) =
%   \cl_{\phi[{\cF_{j-1}}]}(S_1) \setminus
%   \cl_{\phi[{\cF_{j-1}}]}(S_2)$. Thus the induced path labeling
%   $\cl_{\phi[{\cF_{j}}]} = \cl_{j}$. Therefore lemma is proved.  \qed
% \end{proof}

\begin{lemma}
  \label{lem:invar1} In Algorithm~\ref{perms}, at the end of $j$th
  iteration, $j \ge 0$, of the {\bf while} loop, the following
  invariants are maintained.
%   \begin{enumerate}[I {\ }] %\vspace{\topshrink}
%   \item $\cl_j(R)$ is a path in $T$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
%     \ \ \ \ \ \ \ \ \ \ for all $R \in \cF_j$%\vspace{\topshrink}
%   \item $|R| = |\cl_j(R)|$, \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \
%     \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ for all $R \in
%     \cF_j$%\vspace{\topshrink}
%   \item $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, \ \ \ \ \ \ \ \ \ \
%     \ \ \ \ \ \ \ \ \ \ for all $R, R' \in \cF_j$%\vspace{\topshrink}
%   \item $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
%     \cl_j(R'')|$, \ \ \ for all $R, R', R'' \in \cF_j$
%   \end{enumerate}

  \begin{tabular}[h]{lll}
    I & $\cl_j(R)$ is a path in $T$, & for all $R \in \cF_j$\\ 
    II  & $|R| = |\cl_j(R)|$, & for all $R \in \cF_j$ \\ 
    III  & $|R \cap R'| = |\cl_j(R) \cap \cl_j(R')|$, & for all $R, R'
    \in \cF_j$\\ 
    IV  & $|R \cap R' \cap R''|=|\cl_j(R) \cap \cl_j(R') \cap
    \cl_j(R'')|$, & for all $R, R', R'' \in \cF_j$ 
  \end{tabular}
\end{lemma}

% \begin{proof}
%  Proof is by induction on the number of iterations, $j$. In this
%   proof, the term ``new sets'' will refer to the sets added to $\cF_j$
%   in $j$th iteration in line~\ref{setbreak} of Algorithm~\ref{perms},
%   $S_1 \cap S_2, S_1 \setminus S_2, S_2 \setminus S_1$ and its
%   images in $\cl_j$ where $\cl_{j-1}(S_1)$
%   and $\cl_{j-1}(S_2)$ intersect and share a leaf.\\
%   The invariants are true in the base case $(\cF_0, \cl_0)$, since it
%   is the input ICPPL.  Assume the lemma is true till the $j-1$th
%   iteration. Let us consider the possible cases for each of the above invariants for
%   the $j$th iteration.

%   \begin{enumerate}[\textreferencemark]
%   \item {\em Invariant} I/II
%     \begin{enumerate}[{I/II}a $|$] % \textbullet 
%     \item {\em $R$ is not a new set.} It is in $\cF_{j-1}$. Thus
%       trivially true by induction hypothesis.
%     \item {\em $R$ is a new set.} If $R$ is in $\cF_{j}$ and not in
%       $\cF_{j-1}$, then it must be one of the new sets added in
%       $\cF_j$. In this case, it is clear that for each new set, the
%       image under $\cl_j$ is a path since by definition the chosen
%       sets $S_1$, $S_2$ are from $\cF_{j-1}$ and due to the while loop
%       condition, $\cl_{j-1}(S_1)$, $\cl_{j-1}(S_2)$ have a
%       common leaf. Thus invariant I is proved.\\
%       Moreover, due to induction hypothesis of invariant III and the
%       definition of $l_j$ in terms of $l_{j-1}$, invariant II is
%       indeed true in the $j$th iteration for any of the new sets.  If
%       $R = S_1 \cap S_2$, $|R| = |S_1 \cap S_2| = |\cl_{j-1}(S_1) \cap
%       \cl_{j-1}(S_2)| = |\cl_j(S_1 \cap S_2)| = |\cl_j(R)|$.
%       If $R = S_1 \setminus S_2$, $|R| = |S_1 \setminus S_2| = |S_1| -
%       |S_1 \cap S_2| = |\cl_{j-1}(S_1)| - |\cl_{j-1}(S_1) \cap
%       \cl_{j-1}(S_2)| = |\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)| =
%       |\cl_j(S_1 \setminus S_2)|
%       = |\cl_j(R)|$. Similarly if $R = S_2 \setminus S_1$.\\
%     \end{enumerate}
%   \item {\em Invariant} III
%     \begin{enumerate}[{III}a $|$]
%     \item {\em $R$ and $R'$ are not new sets.} It is in
%       $\cF_{j-1}$. Thus trivially true by induction hypothesis.
%     \item {\em Only one, say $R$, is a new set.} Due to invariant IV
%       induction hypothesis, Lemma~\ref{lem:setminuscard} and
%       definition of $\cl_j$, it follows that invariant III is true no
%       matter which of the new sets $R$ is equal to. If $R = S_1 \cap
%       S_2$, $|R \cap R'| = |S_1 \cap S_2 \cap R'| = |\cl_{j-1}(S_1)
%       \cap \cl_{j-1}(S_2) \cap \cl_{j-1}(R')| = |\cl_j(S_1 \cap S_2)
%       \cap \cl_j(R')| = |\cl_j(R) \cap \cl_j(R')|$.  If $R = S_1
%       \setminus S_2$, $|R \cap R'| = |(S_1 \setminus S_2) \cap R'| =
%       |(\cl_{j-1}(S_1) \setminus \cl_{j-1}(S_2)) \cap \cl_{j-1}(R')| =
%       |\cl_{j}(S_1 \cap S_2) \cap \cl_{j}(R')| = |\cl_{j}(R) \cap
%       \cl_{j}(R')|$. Similarly, if $R = S_2 \setminus
%       S_1$. Note $R'$ is not a new set.\\

%     \item {\em $R$ and $R'$ are new sets.} By definition, the new
%       sets and their path images in path label $\cl_j$ are disjoint so
%       $|R \cap R'| = |\cl_j(R) \cap \cl_j(R)| = 0$. Thus case proved.
%     \end{enumerate}
%   \item {\em Invariant} IV

%     Due to the condition in line~\ref{ln:3waycheck}, this invariant is
%     ensured at the end of every iteration.
% %     \begin{enumerate} [{Case 3.}1:]
% %     \item {\em $R$, $R'$ and $R''$ are not new sets.} Trivially
% %       true by induction hypothesis.
% %     \item {\em Only one, say $R$, is a new set.}
% %       If $R = S_1 \cap S_2$,  from Lemma~\ref{lem:fourpaths} and
% %       invariant III hypothesis,  this case is proven. Similarly if $R$
% %       is any of the other new  sets, the case is proven by also using
% %       Lemma ~\ref{lem:setminuscard}.
% %     \item {\em At least two of $R, R', R''$ are new sets.}
% %       The new sets are disjoint hence this case is vacuously true.
% %     \end{enumerate}
%   \end{enumerate} \qed
% %\vspace{-6mm}

% \end{proof}

\begin{lemma}
  \label{lem:noexit1}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{perms} is
  feasible, then the set of hypergraph isomorphism functions that
  defines $(\cF, \cl)$'s feasibility is the same as the set that
  defines $(\cF_j, \cl_j)$'s feasibility, if any.  Secondly, for any
  iteration $j > 0$ of the {\em \bf while} loop, the {\em \bf exit}
  statement in line~\ref{ln:exit1} will not execute.
\end{lemma}
\tnote[TH]{UNCOMMENT}
% \begin{proof}
%   Since $(\cF,\cl)$ is feasible, by Lemma~\ref{lem:feasible}
%   $(\cF_j,\cl_j)$ for every iteration $j > 0$ is feasible.  % Therefore,
% %   every hypergraph isomorphism $\phi: supp(\cF) \rightarrow V(T)$ that
% %   induces $\cl$ on $\cF$ also induces $\cl_{j-1}$ and $\cl_{j}$ on
% %   $\cF_{j-1}$ and $\cF_{j}$ respectively, i.e., $\cl_{\phi[\cF_{j-1}]}
% %   = \cl_{j-1}$ and $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
% %   that for all $x \in supp(\cF)$, for all $v \in V(T)$ the following
% %   hold true.
%   Also, every hypergraph isomorphism $\phi: supp(\cF) \rightarrow
%   V(T)$ that induces $\cl$ on $\cF$ also induces $\cl_{j}$ on
%   $\cF_{j}$, i.e., $\cl_{\phi[\cF_j]} = \cl_j$. Thus it can be seen
%   that for all $x \in supp(\cF)$, for all $v \in V(T)$, if $(x,v) \in
%   \phi$ then $v \in \cl_{j}(S)$ for all $S \in \cF_{j}$ such that $x
%   \in S$.
% % the following
% %   hold true.
% %   \begin{enumerate}[i. ]
% %   \item If $(x,v) \in \phi$ then $v \in \cl_{j-1}(S)$ for all $S \in
% %     \cF_{j-1}$ such that $x \in S$.
% %   \item If $(x,v) \in \phi$ then $v \in \cl_{j}(S)$ for all $S \in
% %     \cF_{j}$ such that $x \in S$
% %   \end{enumerate}
%   In other words, filter 1 outputs a filtered path labeling that
%   ``preserves''
%   hypergraph isomorphisms of the original path labeling.\\
%   Secondly, line~\ref{ln:exit1} will execute \iff the exit condition in
%   line~\ref{ln:3waycheck}, i.e. failure of three way intersection
%   preservation, becomes true in any iteration of the {\em \bf while}
%   loop.  Due to Lemma~\ref{lem:invar1} Invariant IV, the exit
%   condition does not occur if the input is a feasible ICPPL.\qed

% %   such that $\phi(x) = v$ where $v$ is the leaf considered in the
% %   first iterations of while. Clearly, $\phi$ is a renaming of
% %   vertices in hypergraph $\cF$ to those in hypergraph $\cF^\cl$. Thus
% %   the following facts can be observed in every iteration of the loop.

% %   \begin{enumerate}[\hspace{2mm}i. ] \vspace{\topshrink}
% %   \item all intersection cardinalities are preserved in this path
% %     labeling \vspace{\topshrink}
% %   \item element $x$ is exclusive in a hyperedge in $\cF$ since $v$ is
% %     exclusive in a hyperedge in $\cF^\cl$.
% %   \end{enumerate}

% %   Thus the exit condition is never rendered true after $x$ and $v$ are
% %   removed from their respective hyperedges. \qed

% % \noindent
% % This proof uses mathematical induction on the number
% %   of iterations $j$, $j \ge 0$, of the loop that executed
% %   without exiting. The base case, $j = 0$ is obviously true since the
% %   input is an ICPPL and the exit condition cannot hold true due to
% %   ICPPL property (iii).  Assume the algorithm executes till the end
% %   of $j-1$th iteration without exiting at line
% %  ~\ref{ln:3waycheck}. Consider the $j$th iteration. From Lemma
% %  ~\ref{lem:feasible} we know that $(\cF_j, \cl_j)$ and $(\cF_{j-1},
% %   \cl_{j-1})$ are feasible\remove[AS]{and from the proof in lemma
% %     lem:invar1 we know that $(\cF_{j-1}, \cl_{j-1})$ satisfies all the
% %     invariants defined in the lemma}.  Thus there exists a bijection
% %   $\phi: supp(\cF) \rightarrow V(T)$ such that the induced path
% %   % labeling on $\cF_{j-1}$ $\cl_{\phi[\cF_{j-1}]} = \cl_{j-1}$.
% %   labeling on $\cF_{j}$, $\cl_{\phi[\cF_{j}]}$ and on $\cF_{j-1}$,
% %   $\cl_{\phi[\cF_{j-1}]}$ are equal to $\cl_{j}$ and $\cl_{j-1}$
% %   respectively.  We need to prove that for any $R, R', R'' \in
% %   \cF_{j}$, $|R \cap R' \cap R''| = |\cl_j(R) \cap \cl_j(R') \cap
% %   \cl_j(R'')|$.
% %   The following are the possible cases that could arise. From argument
% %   above, $|\cl_j(R) \cap \cl_j(R') \cap \cl_j(R'')| =
% %   |\cl_{\phi[\cF_{j}]}(R) \cap \cl_{\phi[\cF_{j}]} (R') \cap
% %   \cl_{\phi[\cF_{j}]} (R'')|$

% %   \begin{enumerate}[a $|$]
% %   \item {\em None of the sets are new. $R, R', R'' \in \cF_{j-1}$.}
% %     We know $(\cF_{j-1}, \cl_{j-1})$ is feasible. Thus $|R \cap R'
% %     \cap R''| = |\cl_{j-1}(R) \cap \cl_{j-1}(R') \cap \cl_{j-1}(R'')|
% %     = |\cl_{j}(R) \cap \cl_{j}(R') \cap \cl_{j}(R'')|$.
% %   \item {\em Only one, say $R$, is a new set.}  Let $R = S_1 \cap S_2$
% %     ($S_1, S_2$ are defined in the proof of lemma
% %    ~\ref{lem:invar1}). Now we have $|R \cap R' \cap R''| = |S_1 \cap
% %     S_2 \cap R' \cap R''| = |\cl_{j-1}(S_1) \cap \cl_{j-1}(S_2) \cap
% %     \cl_{j-1}(R') \cap \cl_{j-1}(R'')| = |\cl_{j}(R) \cap \cl_{j}(R')
% %     \cap \cl_{j}(R'')|$. Thus proven. If $R$ is any of the other new
% %     sets, the same claim can be verified using lemma
% %    ~\ref{lem:setminuscard}.
% %     % \item []{\bf Case 3:}
% %   \item {\em At least two of $R, R', R''$ are new sets.}  The new sets
% %     are disjoint hence this case is vacuously true.
% %   \end{enumerate}
% %   \qed
% %   \tnote[E2]{remove the induction proof. just text saying x and v are
% %     exclusive in these sets therefore the intersection cardinalities
% %     don't change thus all invariants are still true} 
% \end{proof}

As a result of Algorithm~\ref{perms} each leaf $v$ in $T$
is such that there is exactly one set in $\cF$ with $v$ as a vertex in
the path assigned to it.  In Algorithm~\ref{leafasgn} we identify
elements in $supp(\cF)$ whose images are leaves in a hypergraph
isomorphism if one exists.  Let $S \in \cF$ be such that $\cl(S)$ is a
path with leaf and $v \in V(T)$ is the unique leaf incident on it.  We
define a new path labeling $\cl_{new}$ such that $\cl_{new}(\set{x}) =
\set{v}$ where $x$ an arbitrary element from $S \setminus \bigcup_{\hS
  \ne S} \hS$. In other words, $x$ is an element present in no other
set in $\cF$ except $S$. This is intuitive since $v$ is present in no
other path image under $\cl$ other than $\cl(S)$.  The element $x$ and
leaf $v$ are then removed from the set $S$ and path $\cl(S)$
respectively. After doing this for all leaves in $T$, all path images
in the new path labeling $\cl_{new}$ except leaf labels (a path that
has only a leaf is called the {\em leaf label} for the corresponding
single element hyperedge or set) are paths from a new pruned tree $T_0
= T \setminus \{v \mid v \text{ is a leaf in }
T\}$. Algorithm~\ref{leafasgn} is now presented with details.

\tnote{write a note summarizing filter 1}
% \begin{algorithm}[h]
%   \caption{Leaf labeling from an ICPPL {\tt filter\_2($\cF, \cl, T$)}}
%   \label{leafasgn}
%   \begin{algorithmic}[\lndisplay]
%     \STATE $\cF_0 \assign \cF$, $\cl_0(S) \assign \cl(S)$ for all $S \in \cF_0$
%     \COMMENT {Path images are such that no two path images share a
%       leaf.}
%     \STATE $j \assign 1$\\
%     \WHILE {there is a leaf $v$ in $T$ and a unique $S_1 \in
%       \cF_{j-1}$ such that $v \in \cl_{j-1}(S_1)$ }\label{uniqueleaf}
%     \STATE $\cF_j \assign \cF_{j-1} \setminus \{S_1\}$\\
%     \STATE for all $S \in \cF_{j-1}$ such that $S \ne S_1$ set
%     $\cl_j(S) \assign
%     \cl_{j-1}(S)$\\
%     \STATE $X \assign S_1 \setminus \bigcup_{S \in \cF_{j-1}, S \ne S_1}S$\\
%     \IF{$X$ is empty} \label{xempty} \STATE {\bf exit} \label{ln:exit2} \ENDIF
%     \STATE $x \assign $ arbitrary element from $X$\\
%     \STATE $\cF_j \assign \cF_j \cup \{\{x\}, S_1 \setminus \{x\}\} $\\
%     \STATE $\cl_j(\{x\}) \assign \{v\}$\\
%     \STATE $\cl_j(S_1 \setminus \{x\}) \assign \cl_{j-1}(S_1) \setminus \{v\}$\\
%     \STATE $j \assign j+1$\\
%     \ENDWHILE
%     \STATE $\cF' \assign \cF_j$, $\cl' \assign \cl_j$\\
%     \RETURN $(\cF', \cl')$
%   \end{algorithmic}
% \end{algorithm}

Suppose the input ICPPL $(\cF, \cl)$ is feasible, yet set $X$ in
Algorithm~\ref{leafasgn} is empty in some iteration of the {\bf while}
loop. This will abort our procedure of finding the hypergraph
isomorphism. The following lemma shows that this will not happen.

\begin{lemma}
  \label{lem:xnotempty}
  If the input ICPPL $(\cF, \cl)$ to Algorithm~\ref{leafasgn} is
  feasible, then for all iterations $j > 0$ of the {\em \bf while}
  loop, the {\em \bf exit} statement in line~\ref{ln:exit2} does not
  execute.
\end{lemma}

% \begin{proof}
%   Assume $X$ is empty for some iteration $j > 0$. We know that $v$ is
%   an element of $\cl_{j-1}(S_1)$. Since it is uniquely present in
%   $\cl_{j-1}(S_1)$, it is clear that $v \in \cl_{j-1}(S_1) \setminus
%   \bigcup_{(S \in \cF_{j-1}) \wedge (S \ne S_1)}\cl_{j-1}(S)$.  Note
%   that for any $x \in S_1$ it is contained in at least two sets due to
%   our assumption about cardinality of $X$. Let $S_2 \in \cF_{j-1}$ be
%   another set that contains $x$. From the above argument, we know $v
%   \notin \cl_{j-1}(S_2)$. Therefore there cannot exist a hypergraph
%   isomorphism bijection that maps elements in $S_2$ to those in
%   $\cl_{j-1}(S_2)$. This contradicts our assumption that the input is
%   feasible. Thus $X$ cannot be empty if input is ICPPL and feasible.
%   \qed
% \end{proof}

\begin{lemma}
  \label{lem:invar3}
  In Algorithm~\ref{leafasgn}, for all $j > 0$, at the end of the
  $j$th iteration of the {\bf while} loop the four invariants given in
  Lemma~\ref{lem:invar1} hold.
\end{lemma}

% \begin{proof}
%   By Lemma~\ref{lem:xnotempty} we know that set $X$ will not be empty
%   in any iteration of the {\em \bf while} loop if input ICPPL $(\cF,
%   \cl)$ is feasible and $\cl_j$ is always computed for all $j >
%   0$. Also note that removing a leaf from any path keeps the new path
%   connected. Thus invariant I is obviously true. In every iteration $j
%   > 0$, we remove exactly one element $x$ from one set $S$ in $\cF$
%   and exactly one vertex $v$ which is a leaf from one path
%   $\cl_{j-1}(S)$ in $T$. This is because as seen in
%   Lemma~\ref{lem:xnotempty}, $x$ is exclusive to $S$ and $v$ is
%   exclusive to $\cl_{j-1}(S)$. Due to this fact, it is clear that the
%   intersection cardinality equations do not change, i.e., invariants
%   II, III, IV remain true. On the other hand, if the input ICPPL is
%   not feasible the invariants are vacuously true. \qed
% \end{proof}

%% OLD TEXT TB DELETED?
% \textcolor{cyan}{
% \begin{lemma}
%   \label{lem:notfeasibleexit}
%   \tnote{IS THIS CORRECT?}  If input ICPPL $(\cF, \cl)$ is not
%   feasible, then in one of the recursive calls to Algorithm --3--, the
%   {\em \bf exit} statement in line x in Algorithm --1 or line y in
%   Algorithm --2 will get executed.
% \end{lemma}
% }
%% END OLD TEXT TB DELETED?

We have seen two filtering algorithms above, namely,
Algorithm~\ref{perms} {\tt filter\_1} and Algorithm~\ref{leafasgn}
{\tt filter\_2} which when executed serially respectively result in a
new ICPPL on the same universe $U$ and tree $T$. We also proved that
if the input is indeed feasible, these algorithms do indeed output the
filtered ICPPL. Now we present the algorithmic characterization of a
feasible tree path labeling by way of Algorithm~\ref{al:icppl-find-isomorph}.

Algorithm~\ref{al:icppl-find-isomorph} computes a
hypergraph isomorphism $\phi$ recursively using Algorithm~\ref{perms}
and Algorithm~\ref{leafasgn} and pruning the leaves of the input
tree. In brief, it is done as follows. Algorithm~\ref{leafasgn} gives
us the leaf labels in $\cF_2$, i.e., the elements in $supp(\cF)$ that
map to leaves in $T$, where $(\cF_2, \cl_2)$ is the output of
Algorithm~\ref{leafasgn}. All leaves in $T$ are then pruned away. The
leaf labels are removed from the path labeling $\cl_2$ and the
corresponding elements are removed from the corresponding sets in
$\cF_2$. This tree pruning algorithm is recursively called on the
altered hypergraph $\cF'$, path label $\cl'$ and tree $T'$. The
recursive call returns the bijection $\phi''$ for the rest of the
elements in $supp(\cF)$ which along with the leaf labels $\phi'$ gives
us the hypergraph isomorphism $\phi$.  The following lemma formalizes
the characeterization of feasible path labeling.

\tnote{write a note summarizing filter 1}
% \begin{algorithm}[h]
%   \caption{{\tt get-hypergraph-isomorphism($\cF, \cl, T$)}}
%   \label{al:icppl-find-isomorph}
%   \begin{algorithmic}[\lndisplay]

%     \IF{$T$ is empty}
%     \RETURN $\emptyset$\\
%     \ENDIF
%     \STATE $L \assign \{v \mid v \text{ is a leaf in }      T\}$\\
%     \STATE $(\cF_1, \cl_1) \assign$ {\tt filter\_1($\cF, \cl,
%       T$)}\\
%     \STATE $(\cF_2, \cl_2) \assign$ {\tt filter\_2($\cF_1,
%       \cl_1, T$)}\\

%     \STATE $(\cF', \cl') \assign (\cF_2, \cl_2)$\\
%     \STATE $\phi' \leftarrow \emptyset$

%     \FOR {every $v \in L$}
%     \STATE $\phi'(x) \assign v$ where $x \in \cl_2^{-1}(\{v\})$
%     \COMMENT {Copy the leaf labels to a one to one function $\phi':
%       supp(\cF) \rightarrow L$
%       }\\
%     \STATE Remove $\{x\}$ and $\{v\}$ from $\cF'$, $\cl'$  appropriately\\
%     \ENDFOR

%     \STATE $T' \assign T \setminus L$

%     \STATE $\phi'' \assign$ {\tt
%       get-hypergraph-isomorphism($\cF', \cl', T'$)}
%     \STATE $\phi \assign \phi'' \cup \phi'$ \\
%     \RETURN $\phi$
%   \end{algorithmic}
% \end{algorithm}

\begin{lemma}
  \label{lem:hyperiso}  %{lem:perm}
  If $(\cF, \cl)$ is an ICPPL from a tree $T$ and
  Algorithm~\ref{al:icppl-find-isomorph}, {\tt
    get- hypergraph- isomorphism ($\cF, \cl, T$)} returns a non-empty
  function, then there exists a hypergraph isomorphism $\phi :
  supp(\cF) \rightarrow V(T)$ such that the $\phi$-induced tree path
  labeling is equal to $\cl$ or $\cl_\phi = \cl$.
\end{lemma}
% \begin{proof}
%   It is clear that in the end of every recursive call to
%   Algorithm~\ref{al:icppl-find-isomorph}, the function $\phi'$ is one
%   to one involving all the leaves in the tree passed as input to that
%   recursive call. Moreover, by Lemma~\ref{lem:noexit1} and
%   Lemma~\ref{lem:xnotempty} it is consistent with the tree path
%   labeling $\cl$ passed. The tree pruning is done by only removing
%   leaves in each call to the function and is done till the tree
%   becomes empty. Thus the returned function $\phi: supp(\cF)
%   \rightarrow V(T)$ is a union of mutually exclusive one to one
%   functions exhausting all vertices of the tree. In other words, it is
%   a bijection from $supp(\cF)$ to $V(T)$ inducing the given path
%   labeling $\cl$ and thus a hypergraph isomorphism. \qed
% \end{proof}

\begin{theorem}
  \label{th:charac}
  A path labeling $(\cF, \cl)$ on tree $T$ is feasible \iff it is an
  ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
  T)$ as input returns a non-empty function.
\end{theorem}
% \begin{proof}
%   From Lemma~\ref{lem:hyperiso}, we know that if $(\cF, \cl)$ is an
%   ICPPL and Algorithm~\ref{al:icppl-find-isomorph} with $(\cF, \cl,
%   T)$ as input returns a non-empty function, $(\cF, \cl)$ is feasible.
%   Now consider the case where $(\cF, \cl)$ is feasible. i.e. there
%   exists a hypergraph isomorphism $\phi$ such that $\cl_\phi =
%   \cl$. Lemma~\ref{lem:noexit1} and Lemma~\ref{lem:xnotempty} show us
%   that filter 1 and filter 2 do not exit if input is feasible. Thus
%   Algorithm~\ref{al:icppl-find-isomorph} returns a non-empty
%   function.\qed
% \end{proof}

\subsection*{ICPPL when given tree is a path}
\label{sec:icpplicpia}
Consider a special case of ICPPL with the following properties when
the tree $T$ is a path.  Hence, all path labels are can be viewed as
intervals assigned to the sets in $\cF$.  It is shown, in
\cite{nsnrs09}, that the filtering algorithms outlined above need only
preserve pairwise intersection cardinalities, and higher level
intersection cardinalities are preserved by the Helly Property of
intervals.  Consequently, the filter algorithms do not need to ever
evaluate the additional check to {\em \bf exit}.
%\begin{enumerate}
%\item Given tree $T$ is a path. Hence, all path labels are interval labels.
%\item Only pairwise intersection cardinality
%  preservation is sufficient. i.e. property (iii) in ICPPL is not enforced.
%\item The filter algorithms do not have {\em \bf exit} statements.
%\end{enumerate}
%This is called an Intersection Cardinality Preservation Interval
%Assignment (ICPIA) \cite{nsnrs09}. 
This structure and its algorithm is
used in the next section for finding tree path labeling from a
$k$-subdivided star due to this graph's close relationship with
intervals.

\subsection{Case of a Special Tree -- The $k$-subdivided star}
\label{sec:ksubdivstar}
In this section we consider the problem of assigning paths from a
$k$-subdivided star $T$ to a given set system $\cF$ such that each set
$X \in \cF$ is of cardinality at most $k+2$.  Secondly, we present our
results only for the case when overlap graph $\bO(\cF)$ is connected.
%The
%overlap graph is well-known from the work of
%\cite{kklv10,nsnrs09,wlh02}.  We use the notation in
%\cite{kklv10}. Recall from Section~\ref{sec:prelims} that hyperedges
%$S$ and $S'$ are said to overlap, denoted by $S \overlap S'$, if $S$ and $S'$
%have a non-empty intersection but neither of them is contained in the
%other. The overlap graph $\bO(\cF)$ is a graph in which the vertices
%correspond to the sets in $\cF$, and the vertices corresponding to the
%hyperedges $S$ and $S'$ are adjacent if and only if they overlap.  Note
%that the intersection graph of $\cF$, $\bI(\cF)$ is different from
%$\bO(\cF)$ and $\bO(\cF) \subseteq \bI(\cF)$.  
A connected component of $\bO(\cF)$ is called an overlap component of
$\cF$.  An interesting property of the overlap components is that any
two distinct overlap components, say $\cO_1$ and $\cO_2$, are such
that any two sets $S_1 \in \cO_1$ and $S_2 \in \cO_2$ are disjoint,
or, w.l.o.g, all the sets in $\cO_1$ are contained within one set in
$\cO_2$.  This containment relation naturally determines a
decomposition of the overlap components into rooted containment trees.
We consider the case when there is only one rooted containment tree,
and we first present our algorithm when $\bO(\cF)$ is connected.  It
is easy to see that once the path labeling to the overlap component in
the root of the containment tree is achieved, the path labeling to the
other overlap components in the rooted containment tree is essentially
finding a path labeling when the target tree is a path: each target
path is a path that is allocated to sets in the root overlap
component.  Therefore, for the rest of this section, $\bO(\cF)$ is a
connected graph. Recall that we also consider the special case when
all hyperedges are of cardinality at most $k+2$.  By definition, a
$k$-subdivided star has a central vertex which we call the {\em root},
and each root to leaf path is called a {\em ray}.  First, we observe
that by removing the root $r$ from $T$, we get a collection of $p$
vertex disjoint paths of length $k+1$, $p$ being the number of leaves
in $T$.  We denote the rays by $R_1, \ldots, R_p$, and the number of
vertices in $R_i$, $i \in [p]$ is $k+2$.  Let
$\seq{v_{i1},\ldots,v_{i(k+2)}=r}$ denote the sequence of vertices in
$R_i$, where $v_{i1}$ is the leaf. Note that $r$ is a common vertex to
all $R_i$.

\subsubsection{Description of the Algorithm}
In this section the given hypergraph $\cF$, the $k$-subdivided star
and the root of the star are denoted by $\cO$, $T$ and vertex $r$,
respectively.  In particular, note that the vertices of $\cO$
correspond to the sets in $\cF$, and the edges correspond to the
overlap relation.

%\noindent
For each hyperedge $X \in \cO$, we will maintain a 2-tuple of non-negative
 numbers $\seq{p_1(X), p_2(X)}$.  The numbers satisfy the property that
 $p_1(X) + p_2(X) \leq |X|$, and at the end of path labeling, for each
 $X$, $p_1(X) + p_2(X) = |X|$.  This signifies the algorithm tracking
 the lengths of subpaths of the path assigned to $X$ from at most two
 rays. We also maintain another parameter called the {\em residue} of
 $X$ denoted by $s(X)=|X| - p_1(X)$. This signifies the residue path
 length that must be assigned to $X$ which must be from another
 ray. For instance, if $X$ is labeled a path from only one ray, then
 $p_1(X) = |X|$, $p_2(X) = 0$ and $s(X) = 0$.

 The algorithm proceeds in iterations, and in the $i$-th
 iteration, $i > 1$, a single hyperedge $X$ that overlaps with a
 hyperedge that has been assigned a path is considered.  At the
 beginning of each iteration hyperedges of $\cO$ are classifed into
 the following disjoint sets.
 \begin{enumerate}
 \item [$\cL_1^i$] {\em Labeled without $r$.} Those that have been
   labeled with a path which does not contain $r$ in one of the
   previous iterations.\\  $\cL_1^i = \set{ X \mid p_1(X) = |X| \text{ and
     } p_2(X) = 0 \text{ and } s(X) = 0, X \in \cO}$
 \item [$\cL_2^i$] {\em Labeled with $r$.} Those that have been labeled
   with two subpaths of $\cl(X)$ containing $r$ from two different rays
   in two previous iterations.\\ $\cL_2^i = \set{X \mid 0 < p_1\left(X\right),
     p_2\left(X\right) < |X|=p_1(X)+p_2(X) \text{ and } s(X) = 0, X \in \cO}$
   \item [$\cT_1^i$] {\em Type 1 / partially labeled.} Those that have
     been labeled with one path containing $r$ from a single ray in one
     of the previous iterations. Here, $p_1(X)$ denotes the length of
     the subpath of $\cl(X)$ that $X$ has been so far labeled
     with. \\
     $\cT_1^i = \set{ X \mid 0 < p_1(X) < |X| \text{ and } p_2(X) = 0
       \text{ and } s(X) = |X|-p_1(X), X \in \cO}$
   \item [$\cT_2^i$] {\em Type 2 / not labeled.} Those that have not been
     labeled with a path in any previous iteration.\\
     $\cT_2^i = \set{ X \mid p_1(X) = p_2(X) = 0 \text{ and } s(X) = |X|,
       X \in \cO}$
 \end{enumerate}
% \vspace{-2mm}
% \begin{align*}
%   \cO &= \cL_1^i \cup \cL_2^i \cup \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]\\
%   \cO_i &= \cT_1^i \cup \cT_2^i, \text{    } \forall i \in [p]
% \end{align*}

%\noindent
The set $\cO_i$ refers to the set of hyperedges $\cT_1^i \cup \cT_2^i$
in the $i$th iteration.  Note that $\cO_1 = \cO$.  In the $i$th
iteration, hyperedges from $\cO_i$ are assigned paths from $T$ using
the following rules. Also the end of the iteration, $\cL_1^{i+1},
\cL_2^{i+1}, \cT_1^{i+1}, \cT_2^{i+1}$ are set to $\cL_1^{i},
\cL_2^{i}, \cT_1^{i}, \cT_2^{i}$ respectively, along with some
case-specific changes mentioned in the rules below.

%\noindent
\begin{enumerate}[I.]
\item {\bf Iteration 1:} Let $S=\{X_1,\ldots,X_s\}$ denote the
  super-marginal hyperedges from $\cO_1$.  If $|S|=s \neq p$, then
  exit reporting failure.  Else, assign to each $X_j \in S$, the path
  from $R_j$ such that the path contains the leaf in $R_j$.  This path
  is refered to as $\cl(X_j)$.  Set $p_1(X_j)=|X|, p_2(X_j)=s(X_j)=0$.
  Hyperedges in $S$ are not added to $\cO_2$ but are added to
  $\cL_1^2$ and all other hyperedges are added to
  $\cO_2$.
\item {\bf Iteration $i$:} Let $X$ be a hyperedge from $\cO_i$ such
  that there exists $Y \in \cL_1^i \cup \cL_2^i$ and $X \overlap
  Y$. Further let $Z \in \cL_1^i \cup L_2^i$ such that $Z \overlap Y$.
  If $X \in \cT_2^i$, and if there are multiple $Y$ candidates then
  any $Y$ is selected.  On the other hand, if $X \in \cT_1^i$, then
  $X$ has a partial path assignment, $\cl'(X)$ from a previous
  iteration, say from ray $R_j$. Then, $Y$ is
  chosen such that $X \cap Y$ has a non-empty intersection with a ray
  different from $R_j$.  The key things that are done in assigning a
  path to $X$ are as follows. The {\em end} of path $\cl(Y)$ where
  $\cl(X)$ would overlap is found, and then based on this the
  existence of a feasible assignment is decided.  It is important to
  note that since $X \overlap Y$, $\cl(X) \overlap \cl(Y)$ in any
  feasible assignment.  Therefore, the notion of the {\em end} at
  which $\cl(X)$ and $\cl(Y)$ overlap is unambiguous, since for any
  path, there are two end points.
  \begin{enumerate}
  \item \label{iendpoint} {\em End point of $\cl(Y)$ where $\cl(X)$ overlaps
      depends on $X \cap Z$:} If $X \cap Z
    \neq \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at that
    end of $\cl(Y)$ at which $\cl(Y)$ and $\cl(Z)$ overlap.  If $X
    \cap Z = \emptyset$, then $\cl(X)$ has an overlap of $|X \cap Y|$ at
    that end of $\cl(Y)$ where $\cl(Y)$ and $\cl(Z)$ do not intersect.
  \item {\em Any path of length $s(X)$ at the appropriate end contains
      $r$:} If $X \in \cT_1^i$ then after finding the appropriate end
    as in step~\ref{iendpoint} this the unique path of length $s(X)$
    should end at $r$.  If not, we exit reporting failure.  Else,
    $\cl(X)$ is computed as union of $\cl'(X)$ and this path. If any
    three-way intersection cardinality is violated with this new
    assignment, then exit, reporting failure.  Otherwise, $X$ is added
    to $\cL_2^{i+1}$.  On the other hand, if $X \in \cT_2^i$, then
    after step~\ref{iendpoint}, $\cl(X)$ or $\cl'(X)$ is unique up to
    the root and including it. Clearly, the vertices $\cl(X)$ or
    $\cl'(X)$ contains depends on $|X|$ and $|X \cap Y|$.  If any
    three way intersection cardinality is violated due to this
    assignment, exit, reporting failure.  Otherwise,
    $p_1(X)$ is updated as the length of the assigned path, and $s(X)
    = |X|-p_1(X)$.  If $s(X) > 0$, then $X$ is added to $\cT_1^{i+1}$.
    If $s(X)=0$, then $X$ is added to $\cL_1^{i+1}$.
  \item {\em The unique path of length $s(X)$ overlapping at the
      appropriate end of $Y$ does not contain $r$:} In this case,
    $\cl(X)$ is updated to include this path.  If any three way
    intersection cardinality is violated, exit, reporting failure.
    Otherwise, update $p_1(X)$ and $p_2(X)$ are appropriate, $X$ is
    added to $\cL_1^{i+1}$ or $\cL_2^{i+1}$, as appropriate.
  \end{enumerate}
\end{enumerate}

%\noindent 
\tnote[TH]{UNCOMMENT}
% \subsection*{Proof of Correctness and Analysis of Running Time} 
% It
% is clear that the algorithm runs in polynomial time, as at each step,
% at most three-way intersection cardinalities need to be checked.
% Further, finding super-marginal hyperedges can also be done in
% polynomial time, as it involves considering the overlap regions and
% checking if the inclusion partial order contains a single minimal
% element.  In particular, once the super-marginal edges are identified,
% each iteration involes finding the next hyperedge to consider, and
% testing for a path to that hyperedge.  To identify the next hyperedge
% to consider, we consider the breadth first layering of the hyperedges
% with the zeroeth layer consisting of the super-marginal hyperedges.
% Since $\cO$ is connected, it follows that all hyperedges of $\cO$ will
% be considered by the algorithm.  Once a hyperedge is considered, the
% path to be assigned to it can also be computed in constant time.  In
% particular, in the algorithm the path to be assigned to $X$ depends on
% $\cl(Y), \cl(Z)$, $s(X)$ and the presence or absence of $r$ in the
% candidate partial path $\cl'(X)$.  Therefore, once the super-marginal
% edges are identified, the running time of the algorithm is linear in
% the size of the input.  By the technique used for constructing prime
% matrices \cite{wlh02}, the super-marginal edges can be found in linear
% time in the input size.  Therefore, the algorithm can be implemented
% to run in linear time in the input size.

% %\noindent
% The proof of correctness uses the following main properties:
% \begin{enumerate}
% \item The $k$-subdivided star has a very symmetric structure.  This
%   symmetry is quantified based on the following observation -- either
%   there are no feasible path labelings of $\cO$ using paths from $T$,
%   or there are exactly $p!$ feasible path labelings.  In other words,
%   there is either no feasible assignment, or effectively a unique
%   assignment modulo symmetry.
% \item The $p$ super-marginal hyperedges, if they exist, will each be
%   assigned a path from distinct rays, and each such path contains the
%   leaf.
% \item For a candidate hyperedge $X$, the partial path assignment
%   $\cl'(X)$ is decided by its overlap with $\cl(Y)$ and cardinality of
%   intersection with $\cl(Z)$.
% \end{enumerate}
% These properties are formalized as follows:
% \begin{lemma}
%   \label{lem:sup-mar}
%   If $X \in \cF$ is super-marginal and $\cl$ is a feasible tree path
%   labeling to tree $T$, then $\cl(X)$ will contain a leaf in $T$.
% \end{lemma}
% \begin{proof}
%   Suppose $X \in \cF$ is super-marginal and $(\cF, \cl)$ is a feasible
%   path labeling from $T$.  Assume $\cl(X)$ does not have a leaf.  Let
%   $R_i$ be one of the rays (or the only ray) $\cl(X)$ is part of.
%   Since $X$ is in a connected overlap component, there exists $Y_1 \in
%   \cF$ and $X \nsubseteq Y_1$ such that $Y_1 \overlap X$ and $Y_1$ has
%   at least one vertex closer to the leaf in $R_i$ than any vertex in
%   $X$. Similarly with the same argument there exists $Y_2 \in \cF$
%   with same subset and overlap relation with $X$ except it has has at
%   least one vertex farther away from the leaf in $R_i$ than any vertex
%   in $X$. Clearly $Y_1 \cap X$ and $Y_2 \cap X$ cannot be part of same
%   inclusion chain which contradicts that assumption $X$ is
%   super-marginal. Thus the claim is proved.\qed
% \end{proof}
% \begin{lemma}
%   If $\cO$ does not have any super-marginal edges, then in any
%   feasible path labeling $\cl$ of $\cO$ with paths from $T$ is such
%   that, for any hyperedge $X$ for which $\cl(X)$ contains a leaf, $|X|
%   \geq k+3$.
% \end{lemma}
% \begin{proof}
%   The proof of this lemma is by contradiction.  Let $X$ be a
%   hyperedges such that $|X| \leq k+2$ and that $\cl(X)$ has a leaf.
%   This implies that the overlap regions with $X$, which are captured by
%   the overlap regions with $\cl(X)$, will form a single inclusion
%   chain. This shows that $X$ is a marginal hyperedge which
%   contradicts the assumption that $\cO$ does not have super-marginal
%   hyperedges. \qed
% \end{proof}
% This lemma is used to prove the next lemma for the case when for all
% $X \in \cO$, $|X| \leq k+2$.  The proof is left out as it just uses
% the previous lemma and the fact that the hyperedges in $X$ have at
% most $k+2$ elements.
% \begin{lemma}
%   If there is a feasible path labeling for $\cO$ in $T$, then there
%   are exactly $p$ super-marginal hyperedges.
% \end{lemma}
% These lemmas now are used to prove the following theorem.
% \begin{theorem}
%   Given $\cO$ and a $k$-subdivided star $T$, the above algorithm
%   decides correctly if there is a feasible path labeling $\cl$.
% \end{theorem}
% \begin{proof} {\em Outline.}
% %   If the algorithm outputs a path labeling $\cl$, then it is clear
% %   that it is an ICPPL. The reason is that the algorithm checks that
% %   three-way intersection cardinalities are preserved in each iteration
% %   which ensures \icpplpr~\ref{pr:iii}. Moreover, it is clear that
% %   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
% %   \icpplpr~\ref{pr:i} and \icpplpr~\ref{pr:ii}. For such a labeling
% %   $\cl$, the proof that it is feasible is by induction on $k$. What
% %   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
% %   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
% %   a star. The claim is clear by observing that after Filter 1 and one
% %   iteration of Filter 2, all the leaves have found their pre-images
% %   from support $\cO$.  Therefore, in the induction step, it is clear
% %   that after Filter 1 and one iteration of Filter 2, the leaves are
% %   assigned pre-images.  Removing the leaves from $T$ and the
% %   pre-images from support of $\cO$, results in an ICPPL to a
% %   $(k-1)$-subdivided star.  Now we apply the induction hypothesis, and
% %   we get a isomorphism between the hypergraph $\cO$ and $\cO^\cl$.
%   If the algorithm outputs a path labeling $\cl$, then it is clear
%   that it is an ICPPL. The reason is that the algorithm checks that
%   three-way intersection cardinalities are preserved in each iteration
%   which ensures ICPPL \icpplpr~\ref{pr:iii}. Moreover, it is clear that
%   $\cl(X)$ for any $X \in \cO$ is computed by maintaining
%   ICPPL \icpplpr~\ref{pr:i} and ICPPL \icpplpr~\ref{pr:ii}. For such a labeling
%   $\cl$, the proof that it is feasible is by induction on $k$. What
%   needs to be shown is that Algorithm~\ref{al:icppl-find-isomorph}
%   successfully runs on input $(\cO, \cl)$. In base case $k=0$, $T$ is
%   a star. Also every set is at most size 2 ($k+2$) size and thus
%   overlaps are at most 1. If two paths share a leaf in {\tt filter\_1}
%   one must be of length 2 and the other of length 1. Thus the exit
%   condition is not met. Now it is trivial that the exit condition in
%   {\tt filter\_2} is also not met. Thus claim proven for base case.
%   Now assume the claim to be true when target tree is a
%   $(k-1)$-subdivided star. Consider the case of a $k$-subdivided
%   star. It is clear that after {\tt filter\_1} and one iteration of
%   {\tt filter\_2}, the leaves are assigned pre-images.  Removing the
%   leaves from $T$ and the pre-images from support of $\cO$, results in
%   an ICPPL to a $(k-1)$-subdivided star.  Now we apply the induction
%   hypothesis, and we get a isomorphism between the hypergraph $\cO$
%   and $\cO^\cl$.

% %  \noindent
%   In the reverse direction if there is a feasible path labeling $\cl$,
%   then we know that $\cl$ is unique up to isomorphism.  Therefore,
%   again by induction on $k$ it follows that the algorithm finds $\cl$.
%   \qed
% \end{proof}

\section{Conclusion}
\tnote{MAKE CONNECTION TO PATH GRAPH ISOMORPHISM HERE}
As it will be described in detail later in this document, isomorphism
of certain classes of graphs, namely chordal graphs, have a close
relationship with consecutive ones property and generalizations of it.
This is perhaps because of how closely COP of a matrix relates to
properties of graphs derived from matrices as seen in the following
results. %-- Other relations of COP with Graphs: \\
\tnote[TH]{peo exists iff chordal. lexicographic BFS
  [tag:chordalGraph]} 
A well known result in perfect graph theory is that the maximal
cliques of an interval graph $G$ can be linearly ordered such that for
all $v \in V(G)$, cliques containing $v$ are consecutive in the

ordering\cite{gh64}. This clearly means that a graph $G$ is an
interval graph if and only if \tnote{verify from paper the
  statement of claim.} maximal clique vertex incidence matrix of $G$
has COP.  Also, maximal cliques of any chordal graph can be enumerated
in polytime $O(m+n)$ \tnote{citation?!!}.  \cite{fg65} uses these
results to give the first polynomial time algorithm for
COT. \tnote[TH]{check. how do they use it?}  A bipartite graph
is convex (on $R$) if and only if its half adjacency matrix has COP on
rows.  The results in \cite{bl76}
%\cite{bl62} 
on COT are based on the result that
interval graphs are AT-free chordal graphs. \tnote{the latter being
  Tucker's?}

\tnote[TH]{How COP (and TPL) relates to interval graph
  isomorphism (and path graph isomorphism) move/summarize what is
  already in Motivation section.}

\tnote[TH]{TBD survey -- see blue notes under Graph
  Isomorphism. namely citations in cite:aas93}

\tnote[TH]{a brief on heirarchy of chordal graphs, path
  graphs, interval graphs, peo, clique tree etc. and results.}

\tnote[TH]{interval graphs are incomparability graphs. see
  Golumbic. [tag:classification]}

\tnote[TH]{finding min length hole in bipartite graph is
  polynomial time [Sec 3.3.2 in cite:d08phd]}

\tnote[TH]{theorem 2.2 in cite:d08phd - $G$ is union of
  v.d. caterpillars iff $M$ has COP in rows. $M$ is edge vertex
  incidence matrix $(*,2)$ of $G$}

{\bf Complexity challenges:} Canonization is an important tool in graph isomorphism. Invention of a
deterministic method of canonization for any class of graphs naturally
results in an algorithm for isomorphism. All that is required is to
check if the canons of two graphs are the same. Thus complexity of
graph isomorphism can be studied by studying canonization
methods. While general graph ismorphism remains elusive in terms of
hardness, canonization has been studied for smaller classes of graphs
thus giving complexity/hardness results for them.

In 1992, \cite{sl92} made an important discovery that tree isomorphism
is in logspace. It was done by inventing a method of canonization of
trees using a logspace depth first traversal
algorithm. \tnote{confirm the statement of claim}.

\cite{kklv10} proved that interval graph canonization is also in
logspace thus drawing logspace conslusions about COT. 
% This result is
% built on the fact that 
Interval graphs are FP+C (fixed point with
counting) definable \tnote{elucidate} and \cite{l10} showed that
this implies that it captures PTIME. \tnote{elucidate} This result
along with that of undirected graph reachability being in logspace
\cite{rei08}, \cite{kklv10} proved their logspace result.\\

\tnote[TH]{the interesting survey in conclusion section
  of kklv10}

\tnote[TH]{[[[COP helping some problems with hardness]]] having
  COP in the input structure makes some problems less hard than in a
  general input. -- min set cover problem is known to be NP hard. so
  is min hitting
  set. these two problems are equivalent cite-adp80\\
  -- -- complexity results: log factor polynomial time approximable cite-f98, W[2]-complete with solution size as parameter cite-df99\\
  -- a couple of generalizations:\\
  -- -- min degree hypergraph (MDH) problem\\
  -- -- red blue covering (RBSC) problem\\
----- cite-d08phd Chapter 5.\\
----- what are cite-d08phd results\\
----- Rectangle stabbing \\
----- see cite-d08phd Chapter 6?
}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography.
\begin{singlespace}
%\begin{thebibliography}{10}
\addcontentsline{toc}{section}{Bibliography}

{\footnotesize
  \bibliographystyle{alpha} 
  \bibliography{../lib/cop-variants__thesis}
}
%\end{thebibliography}
\end{singlespace}
\newpage

%%%%%%

%\chapter*{Proposed Contents of the Thesis}
\section{Proposed Contents of the Thesis}
%\addcontentsline{toc}{chapter}{Proposed Contents of the Thesis}
The outline of the thesis is as follows:

\begin{enumerate}[{Chapter} 1]
\item - Introduction
  \begin{enumerate} [{Section 1.}1]
  \item Consecutive Ones Testing
  \item Matrix modification to attain COP or ``almost'' COP
  \item Graph Isomorphism
  \item Logspace complexity of graph canonization using COP
  \item Motivation, Objective and Scope of Thesis
  \item Summary of Results
  \item Organization of document
  \end{enumerate}

\item - Consecutive ones property % ( survey of cop includes mainly *
  % dom's thesis (perhaps minus chap 4?) * NSNSR09 * McConnel's paper
  % *
  % KKLV10 to name a few. sections will be as follows)\\
  \begin{enumerate}[{Section 2.}1]
  \item Characterization of COP
  \item Recognition of COP
  \item Alteration to matrices to get COP
  \item Complexity of certain COP variations
  \end{enumerate}

\item - Other problems related to COP %(see literature)

\item Research %- Extension to trees
  \begin{enumerate}[{Section 4.}1]
  \item Tree path labeling of path
    hypergraphs%: ( stuff in latin paper,)
  \item Extension of po theory from \cite{nsnrs09}
    % (the matrix theory submitted in walcom, any other work - need to
    % see notes.)
  \end{enumerate}

\item - Conclusion

\item - Bibliography

\end{enumerate}

%\vskip 4cm
% \chapter{PUBLICATIONS}
% \section{Papers in Refereed Journals}
% \begin{enumerate}
% \item Title \\
% 	{\bf Author 1}, Author 2...\\
% 	{\it Journal title.}, {\bf Volume}, Page (Year)
% \end{enumerate}

% \section{Presentations in Conferences}
% \begin{enumerate}
% \item  Presented titled  {\em Title} at the {\bf Conference on...}  place, date..
% \end{enumerate}

\end{document}