### 2012-01-03 src/001certificate.tex
Cleaned up thesis.tex. Split thesis.tex into chapterwise subfiles for easy edit tex management. Content changes are mainly in chapter 1 and 3 although nothing significant. To being working on chapter 2 after this commit.

\certificate

\vspace*{0.5in}

\noindent 
This is to certify that the thesis titled {\bf
  \mythesistitleNOTSYNPOSIS}, submitted by {\bf \myname}, to the {\bf
  Indian Institute of Technology Madras}, for the award of the degree
of {\bf \mydegree}, is a bona fide record of the research work done by
her under our supervision.  The contents of this thesis, in full or in
parts, have not been submitted to any other Institute or University
for the award of any degree or diploma.

\vspace*{1.5in}

\begin{singlespacing}
  \hspace*{-0.25in}
  \parbox{3in}{
    \noindent {\bf Dr.~\myadvisor} \\
    \small
    \noindent Research Guide\\
    \noindent Associate Professor\\
    \noindent Dept. of \mydept\\
    \noindent IIT Madras -- 600 036 } \hfill
  \parbox{1.25in}{               % {1.75in}
    \small
    \vspace*{0.75in}
    \begin{tabular}[h]{r}
    \noindent {Chennai}\\         
    \noindent {\mythesissubmissiondate} 
    \end{tabular}
  }
\end{singlespacing}