### 2012-01-03 src/thesis.tex

\documentclass[MS]             % MS thesis dissertation.
              {iitmdiss_as}    % Hacked by hyperbolicme.
%              {iitmdiss}      % If the powers to be are finicky.

\usepackage{times}             % Font?
\usepackage{t1enc}             % T1 encoding??  
\usepackage[T1,OT1]{fontenc}
\usepackage{listings}

\usepackage{../lib/mythesislib}
\usepackage{../lib/mylatexlib}
% \usepackage{comment}         % DEBUG ONLY
\includeonly{010chap1,020chap2}
%\includeonly{0021ded}

\def \mythesissubmissiondate {10 January 2012}
\def \mythesissubmissionmonth {\MakeUppercase {January 2012}}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title page
%
\title{\mythesistitleNOTSYNPOSIS} \author{\myname} \date{
  {\mythesissubmissionmonth}}
\department{\MakeUppercase{\mythesisdept}}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Certificate
%
\include{001certificate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Acknowledgements
%
\include{002ack}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dedication?
%
\include{0021ded}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abstract
%
\include{003abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Table of contents etc.
%
\begin{singlespace}
  \tableofcontents
  \thispagestyle{empty}
  \listoftables \addcontentsline{toc}{chapter}{LIST OF TABLES}
  \listoffigures \addcontentsline{toc}{chapter}{LIST OF FIGURES}
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Abbreviations
%
\include{004abbrv}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Notation
%
\include{005notation}
%\pagebreak \clearpage

\pagenumbering{arabic} % The main text will follow from this point so
                       %  set the page numbering to arabic from here on.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Main body Chapters
%
\include{010chap1}
\include{020chap2}
\include{030chap3}
\include{040chap4}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Appendices.
%
% \appendix %%% NON NEIN NAHI ILLA NONE
% \chapter{A}
% \newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography.
%
\begin{singlespace}
  \addcontentsline{toc}{chapter}{REFERENCES \tnote[minor]{make names
    in bib file uniform style w.r.t. firstname/initials}}

  {\footnotesize%
    \bibliographystyle{alpha}%
    \bibliography{../lib/cop-variants__thesis} %
  }
\end{singlespace}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% List of papers
%
% \listofpapers %%% NON NEIN NAHI ILLA NONE
% \begin{enumerate}
% \item Authors....  \newblock Title...  \newblock {\em Journal},
%   Volume, Page, (year).
% \end{enumerate}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% DOCUMENTATION ON THE THESIS CLASS FILE ETC.

% Before reading any further please note that you are strongly advised
% against changing any of the formatting options used in the class
% provided in this directory, unless you are absolutely sure that it
% does not violate the IITM formatting guidelines.  \emph{Please do not
%   change the margins or the spacing.}  If you do change the formatting
% you are on your own (don't blame me if you need to reprint your entire
% thesis).  In the case that you do change the formatting despite these
% warnings, the least I ask is that you do not redistribute your style
% files to your friends (or enemies).

% It is also a good idea to take a quick look at the formatting
% guidelines.  Your office or advisor should have a copy.  If they
% don't, pester them, they really should have the formatting guidelines
% readily available somewhere.

% To generate PDF's with the links from the \verb+hyperref+ package use
% the following command:
% \begin{verbatim}
% % dvipdfm -o thesis.pdf thesis.dvi
% \end{verbatim}

% \section{Package Options}

% Use this thesis as a basic template to format your thesis.  The
% \verb+iitmdiss+ class can be used by simply using something like this:
% \begin{verbatim}
% \documentclass[PhD]{iitmdiss}  
% \end{verbatim}

% To change the title page for different degrees just change the option
% from \verb+PhD+ to one of \verb+MS+, \verb+MTech+ or \verb+BTech+.
% The dual degree pages are not supported yet but should be quite easy
% to add.  The title page formatting really depends on how large or
% small your thesis title is.  Consequently it might require some hand
% tuning.  Edit your version of \verb+iitmdiss.cls+ suitably to do this.
% I recommend that this be done once your title is final.

% To write a synopsis simply use the \verb+synopsis.tex+ file as a
% simple template.  The synopsis option turns this on and can be used as
% shown below.
% \begin{verbatim}
% \documentclass[PhD,synopsis]{iitmdiss}                                
% \end{verbatim}

% Once again the title page may require some small amount of fine
% tuning.  This is again easily done by editing the class file.

% This sample file uses the \verb+hyperref+ package that makes all
% labels and references clickable in both the generated DVI and PDF
% files.  These are very useful when reading the document online and do
% not affect the output when the files are printed.

% \section{Example Figures and tables}

% Fig.~\ref{fig:iitm} shows a simple figure for illustration along with
% a long caption.  The formatting of the caption text is automatically
% single spaced and indented.  Table~\ref{tab:sample} shows a sample
% table with the caption placed correctly.  The caption for this should
% always be placed before the table as shown in the example.

% \begin{figure}[htpb]
%   \begin{center}
%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
%     \resizebox{50mm}{!} {\includegraphics *{iitm.eps}}
%     \caption {Two IITM logos in a row.  This is also an
%       illustration of a very long figure caption that wraps around two
%       two lines.  Notice that the caption is single-spaced.}
%   \label{fig:iitm}
%   \end{center}
% \end{figure}

% \begin{table}[htbp]
%   \caption{A sample table with a table caption placed
%     appropriately. This caption is also very long and is
%     single-spaced.  Also notice how the text is aligned.}
%   \begin{center}
%   \begin{tabular}[c]{|c|r|} \hline
%     $x$ & $x^2$ \\ \hline
%     1  &  1   \\
%     2  &  4  \\
%     3  &  9  \\
%     4  &  16  \\
%     5  &  25  \\
%     6  &  36  \\
%     7  &  49  \\
%     8  &  64  \\ \hline
%   \end{tabular}
%   \label{tab:sample}
%   \end{center}
% \end{table}

% \section{Bibliography with BIB\TeX}

% I strongly recommend that you use BIB\TeX\ to automatically generate
% your bibliography.  It makes managing your references much easier.  It
% is an excellent way to organize your references and reuse them.  You
% can use one set of entries for your references and cite them in your
% thesis, papers and reports.  If you haven't used it anytime before
% please invest some time learning how to use it.

% I've included a simple example BIB\TeX\ file along in this directory
% called \verb+refs.bib+.  The \verb+iitmdiss.cls+ class package which
% is used in this thesis and for the synopsis uses the \verb+natbib+
% package to format the references along with a customized bibliography
% style provided as the \verb+iitm.bst+ file in the directory containing
% \verb+thesis.tex+.  Documentation for the \verb+natbib+ package should
% be available in your distribution of \LaTeX.  Basically, to cite the
% author along with the author name and year use \verb+\cite{key}+ where
% \verb+key+ is the citation key for your bibliography entry.  You can
% also use \verb+\citet{key}+ to get the same effect.  To make the
% citation without the author name in the main text but inside the
% parenthesis use \verb+\citep{key}+.  The following paragraph shows how
% citations can be used in text effectively.

% More information on BIB\TeX\ is available in the book by
% \cite{lamport:86}.  There are many
% references~\citep{lamport:86,prabhu:xx} that explain how to use
% BIB\TeX.  Read the \verb+natbib+ package documentation for more
% details on how to cite things differently.

% Here are other references for example.  \citet{viz:mayavi} presents a
% Python based visualization system called MayaVi in a conference paper.
% \citet{pan:pr:flat-fst} illustrates a journal article with multiple
% authors.  Python~\citep{py:python} is a programming language and is
% cited here to show how to cite something that is best identified with
% a URL.

% \section{Other useful \LaTeX\ packages}

% The following packages might be useful when writing your thesis.

% \begin{itemize}  
% \item It is very useful to include line numbers in your document.
%   That way, it is very easy for people to suggest corrections to your
%   text.  I recommend the use of the \texttt{lineno} package for this
%   purpose.  This is not a standard package but can be obtained on the
%   internet.  The directory containing this file should contain a
%   lineno directory that includes the package along with documentation
%   for it.

% \item The \texttt{listings} package should be available with your
%   distribution of \LaTeX.  This package is very useful when one needs
%   to list source code or pseudo-code.

% \item For special figure captions the \texttt{ccaption} package may be
%   useful.  This is specially useful if one has a figure that spans
%   more than two pages and you need to use the same figure number.

% \item The notation page can be entered manually or automatically
%   generated using the \texttt{nomencl} package.

% \end{itemize}

% More details on how to use these specific packages are available along
% with the documentation of the respective packages.

% END OF DOCUMENTATION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%