### 2012-01-06 src/0021ded.tex
Added more stuff to chapter 2. see diff.

%\vskip 15ex%
\vspace*{1in}

{
  \fontencoding{T1}
%  \fontfamily{cmdh} 
  \fontfamily{bch} %{put}  %{bch} % {pbk}
%  \fontfamily{pag} %
  \selectfont%
  {\large \noindent The process of preparing programs for the digital
    computer is especially attractive, not only because it can be
    economically and scientifically rewarding, but also because it can
    be an aesthetic experience much like composing poetry or
    music.%
    {\small \par \hfill \emph{The Art of Computer Programming},
      D.\;E.\;K.}%
  }%
  \vskip 5ex%
  {\large \noindent The rhetoric and mythos of science create the
    comforting image of linear progression toward truth.%
    {\normalsize \par \hfill \emph{New Oxford American
        Dictionary, 2nd Ed.}}%\textbf{mythos};
  }%  
  \vskip 15ex%
  {\large%
    \noindent Art is the solution to chaos. \par %
    {\normalsize \par \hfill \emph{citation needed.}}
  }%
}