### 2012-01-07 src/0021ded.tex
chapter 2 - started set system side of things. see diff for more details.

%\vskip 15ex%
\vspace*{1in}

{
  \fontencoding{T1}
%  \fontfamily{cmdh} 
  \fontfamily{bch} %{put}  %{bch} % {pbk}
%  \fontfamily{pag} %
  \selectfont%
  {\large \noindent The process of preparing programs for the digital
    computer is especially attractive, not only because it can be
    economically and scientifically rewarding, but also because it can
    be an aesthetic experience much like composing poetry or
    music.%
    {\small \par \hfill \emph{The Art of Computer Programming},
      D.\;E.\;K.}%
  }%
  \vskip 5ex%
  {\large \noindent The rhetoric and mythos of science create the
    comforting image of linear progression toward truth.%
    {\normalsize \par \hfill \emph{New Oxford American
        Dictionary, 2nd Ed.}}%\textbf{mythos};
  }%  
%   \vskip 5ex%
%   {\large \noindent Art is the triumph over chaos.%
%     %{\normalsize \par \hfill \emph{JC}}%
%   }%
%   \vskip 5ex%
%   {\large \noindent Art, in itself, is an attempt to bring order out of chaos.%
%     %{\normalsize \par \hfill \emph{SS}%
%   }%
  \vskip 5ex%
  {\large%
    \noindent \emph{L'art c'est la solution au chaos.}%(Art is the solution to chaos.) %
    %{\normalsize \par \hfill \emph{citation needed.}}
  }%
}