### 2012-01-14 src/0031ded.tex
Started cleaning up chapter 3. Deferring it though to work on Chapter 2. Need to accommodate N's new proofs for removal of algorithmic charac.
See diff for other changes left out in this comment.

\definecolor{dim}{rgb}{0.7,0.7,0.7}
\def \dimspace {0.125in}

{%
  \fontencoding{T1}%
  \fontfamily{bch} %{put} %{bch} % {pbk} % \fontfamily{cmdh}
  % \fontfamily{pag}
  \selectfont%

  \vglue -0.5in % this makes top margin 2in
%   \vskip -5ex%
%   \hspace{5in}\parbox{1in}{\raggedright Hello world}

  \hspace{-1in}%
  \parbox[c]{4in}{%
    \large \raggedright \noindent %
    \color{dim}{%
      The process of preparing programs for the digital computer is
      especially attractive, not only because it can be economically
      and scientifically rewarding, but also because it can be an
      aesthetic experience much like composing poetry or music.%
      {%
        \small \par \hfill \emph{The Art of Computer Programming},
        D.\,E.\,K.%
      }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{1.5in}%
  \parbox[t]{4in}{%
    \Large \noindent \raggedleft%
    \color{dim}{%
      The rhetoric and mythos of science create the comforting image
      of linear progression toward truth.%
      {%
        \small \par \hfill%
        \textbf{mythos}; \emph{New Oxford American Dictionary, 2nd
          Ed.}%
      }%
    }%
  }
  \\
  \vspace{\dimspace}

  % \hspace{1in}
  {\centering
    \parbox{6in}{%
%      \Huge 
      \fontsize{27}{1.5}\selectfont
      \centering \noindent %
%      \textbf{
      \emph{L'art c'est la solution au chaos.} \par \noindent
      \noindent (Art is the solution to chaos.)
%      }
%       {%
%         \small \par \hfill%
%         \emph{citation needed.}%
%       }%
    }%
  }%

  \vspace{\dimspace}

  \hspace{-0.25in}%
  \parbox{2.5in}{%
    \Large 
    \raggedright \noindent%
    \color{dim}{%
      Kilroy wasn't Haar.  Free the group.  Nuke the kernel.  Power to
      the $n$.  $N = 1 \Rightarrow P = NP$ %
    }%
    {%
      \small \par \hfill%
      \emph{Concrete Mathematics}%
    }%
  }%
  \vspace{\dimspace}

  \hspace{1in}%
  \parbox{5in}{%
    \large \raggedleft \noindent%
    \color{dim}{%
      Anything that happens, happens.  Anything that, in happening,
      causes something else to happen, causes something else to
      happen.  Anything that, in happening, causes itself to happen
      again, happens again.  It doesn't necessarily do it in
      chronological order, though.  }%
    {%
      \small \par \hfill%
      \emph{The Salmon of Doubt}, D.\,N.\,A.%
    }%
  }%
  \vspace{\dimspace}

  \hspace{-0.50in}%
  \parbox{4in}{%
    \huge \noindent \raggedright%
    \color{dim}{%
      This is water, this is water.%
      {%
        \small \par \hfill%
        D.\,F.\,W.%
      }%
    }%
  }%

}