### 2012-01-18 src/_region_.tex

\message{ !name(020chap2.tex)}
\message{ !name(020chap2.tex) !offset(707) }
  \item The relative complements $A \setminus B$ and $B \setminus A$
    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
  \item Also note that trivially, sets in $\cT(U)$ are consecutive in
    any permutation of $U$\footnote{$\emptyset$ is considered
      consecutive by convention.}.
\end{enumerate}
\end{observation}

\begin{theorem}[{\cite[Th~3]{mm96}}]
  For any set system $\cF$, the set of all \COP permutations of $\cF$
  is equal to the set of all \COP permutations of $\overline{\cF}$.
\end{theorem}

\message{ !name(020chap2.tex) !offset(-17) }