### 2012-02-02 src/0031ded.tex
-- sent to n for review
- moved ded to quot. added ded.
- pqtree, pctree figures completed.
- fixed a few missing bibs
- random clean up including temptexts. see diff.

\xclearpage

{% 
  \thispagestyle{empty}%
  \parbox[c]{\linewidth}%
  {%
    \vspace{0.5\linewidth}% 
    \centering% 
    % To you, Z., for your patience.%
    \large%
    {\em For Z.}%
  }%
}%