### 2012-02-24 src/020chap2.tex

\xclearpage

%\chapter{\CoP\,-- A Survey of Important Results}
\chapter[COP -- A Survey]{Consecutive-ones Property -- A Survey of
  Important Results}
\label{ch:copsurvey}

This chapter surveys several results that are significant to this
thesis or to \COP in general. These predominantly pertain to
characterizations of \COP, algorithmic tests to check for \COP,
optimization problems on binary matrices that do not have \COP and
some applications of \COP.

The chapter is organized as follows. Section~\ref{sec:copgraphtheory}
discusses the close connections of matrices with \COP in graph
theory. Section~\ref{sec:surveycoptest} discusses the important
results in the area of \cop testing (COT) -- we describe the history
of this problem's several solutions including Tucker's forbidden
submatrices, \PQtree, \PQRtree and ICPIA.

It is worth noting that we make a very interesting observation about
\PQRtree, generalised \PQtree and \gPQtree\ -- that the theory that
all these independent data structures are built upon are very similar
and equivalent. This is exhaustively described in
Section~\ref{sec:surveycertalgo} and curiously, this comparison has
not been encountered by us in the literature.

Section~\ref{sec:surveycopopt} briefly touches upon the scenario of
algorithmically verifying that a matrix indeed does not have \COP. Its
significance lies in the fact that almost all \COT algorithms, upon
detection of no \COP, simply reports the same without any certificate
for verification during algorithm implementation. The two approaches
presented in this section gives a certificate.

\section{\COP in Graph Theory}
\label{sec:copgraphtheory}

\COP is closely connected to several types of graphs by way of
describing certain combinatorial graph properties. There are also
certain graphs, like convex bipartite graphs, that are defined solely
by one of its associated matrix having \COP.  In this section we will
see the relevance of \cop to graphs.  To see this we introduce certain
binary matrices that are used to define graphs in different
ways. While adjacency matrix is perhaps the most commonly used such
matrix, Definition~\ref{def:graphmatrices} defines this and a few
more.

\begin{definition}
  {\em Matrices that define graphs. \cite[Def.~2.4]{d08phd}.}  
  \label{def:graphmatrices} 
  Let $G$ and $H$ be defined as follows. $G = (V,E_G)$ is a graph with
  vertex set $V = \{v_i \mid i \in [n]\}$ and edge set $E_G \subseteq
  \{(v_i,v_j) \mid i, j \in [n]\}$ such that $|E_G| = m$. $H = (A, B,
  E_H)$ is a bipartite graph with partitions $A = \{a_i \mid i \in
  [n_a]\}$ and $B = \{b_i \mid i \in [n_b]\}$.
  \begin{enumerate}[{\ref{def:graphmatrices}}--i.]
  \item \emph{Adjacency matrix} of $G$ is the symmetric $n \times n$
    binary matrix $M$ with $m_{i,j} = \un$ \iff $(v_i,v_j) \in E_G$
    for all $i,j \in [n]$.
  \item \emph{Augmented adjacency matrix} of $G$ is obtained from its
    adjacency matrix by setting all main diagonal elements to \un, \ie
    $m_{i,i} = \un$ for all $i \in [n]$.
  \item \emph{Maximal clique matrix} or \emph{vertex-clique incidence
      matrix} of $G$ is the $n \times k$ binary matrix $M$ with
    $m_{i,j} = \un$ \iff $v_i \in C_j$ for all $i \in [n], j \in [k]$
    where $\{C_j \mid j \in [k]\}$ is the set of maximal cliques of
    $G$.
    \label{def::maxcliquematrix}
  \item \emph{Half adjacency matrix} of $H$ is the $n_a \times n_b$
    binary matrix $M$ with $m_{i,j} = \un$ \iff $(a_i, b_j) \in E_H$.
  \end{enumerate}
\end{definition}

\figgraphmatrices

Now we will see in Definition~\ref{def:graphwithcop} certain graph
classes that is related to \COP or \CROP.\\

\begin{definition}{\emph{Graphs that relate to
      COP.\cite[Def.~2.5]{d08phd}}} %
  \label{def:graphwithcop} %
  Let $G$ be a graph and $H$ be a bipartite graph.
  \begin{enumerate}[\ref{def:graphwithcop}--i.]
  \item $G$ is \emph{convex-round} if its adjacency matrix has the
    \CROP.
  \item \label{def::concave-round} $G$ is \emph{concave-round} if its
    augmented adjacency matrix has \CROP. \cite{bhy00}
  \item $G$ is an \emph{interval graph} if its vertices can be mapped
    to intervals on the real line \stt two vertices are adjacent \iff
    their corresponding intervals overlap \cite{sb59}.  $G$ is an interval graph \iff its maximal clique matrix
    has COP \cite{fg65} (This follows \cite{gh64} which states
      that the maximal cliques of interval graph $G$ can be linearly
      ordered \stt for all $v \in V(G)$, cliques containing $v$ are
      consecutive in the ordering \cite[Th. 8.1]{mcg04}.)
    \begin{enumerate}[a.]
    \item $G$ is a \emph{unit interval graph} if it is an interval
      graph \stt all intervals have the same length.
    \item $G$ is a \emph{proper interval graph} if it is an interval
      graph \stt no interval properly contains another.
    \end{enumerate}
    The set of unit interval graphs and the set of proper interval
    graphs coincide \cite{rob69, gar07}.
 \item $G$ is a \emph{circular-arc graph} if its vertices can be
    mapped to a set of arcs on a circle \stt two vertices are adjacent
    \iff their corresponding arcs overlap.
  \item $H$ is \emph{convex bipartite on columns (rows)} if its half
    adjacency matrix has \COP on rows (columns).%
    \label{def::convexbi}
  \item $H$ is \emph{biconvex bipartite} or \emph{doubly
      convex}\cite{yc95} if its half adjacency matrix has \COP on both
    rows and columns.
  \item $H$ is \emph{circular convex} if its half adjacency matrix has
    \CROP.
  \end{enumerate}
\end{definition}

Interval graphs\footnote{\cite{mcc04} cites that the problem of
  recognizing interval graphs has significance in molecular
  biology. Interestingly, in the late 1950s, before the structure of
  DNA was well-understood, Seymour Benzer was able to show that the
  intersection graph of a large number of fragments of genetic
  material was an interval graph \cite{sb59}. This was regarded as
  compelling evidence that genetic information was somehow arranged
  inside a structure that had a linear topology which we now know to
  be true from the discovery of linear structure of DNA.}and
circular-arc graphs have a long history in research.  The interest
around them is due to their very desirable property that several
problems that are NP-complete on general graphs, like finding a
maximum clique or minimum coloring or independent set, are polynomial
time solvable in these graph classes \cite{clrs01}.  In a similar
fashion, a lot of problems that are hard on general matrices have
efficient solutions on matrices with \COP or \CROP \cite[more
citations pg.\,33]{d08phd}.

Table~\ref{tab:graphmatrices} summarises the way these graphs are
characterized by their matrices having \COP or \CROP.
Our focus in this chapter (and thesis) is mainly \COP and 
having seen how useful \COP is in identifying or characterizing many
types of graphs, we will now see results that study recognition of \COP
in matrices in the following section.

\tabgraphmatrices

\section{Matrices with COP}
\label{sec:surveycoptest}

The most important questions with respect to a particular 
property desired in a structure/object are perhaps the following.
\begin{itemize}%[]
\singlespacing
\item Does the desired property exist in the given input?
\item If the test is affirmative, what is a certificate of the affirmative?
\item If the test is negative, what are the optimization possibilities
  for the property in the input? In other words, how close to having the
  property can the input be?
\item If the test is negative, what is a certificate of the negative?
\end{itemize}

In this section and the rest of the chapter we see results that shaped the
correponding areas respectively for \cop in binary matrices.

\begin{enumerate}[a.]
\singlespacing
\item \label{q:testcop} Does a given binary matrix have \COP?
\item \label{q:copperm} What is the \COP permutation for the given matrix with \COP?
\item \label{q:copopt} What are the optimizations possible and practically useful on
  the given matrix without \COP?
\item \label{q:nocopcert} If algorithm for (\ref{q:testcop}) returns
  \textbf{false}, can a certificate for this be computed?
\end{enumerate}

Without doubt, besides computing answers to these questions, we are
interested in the efficiency of these computations in terms of
computational complexity theory. Results towards
questions~(\ref{q:testcop}) and (\ref{q:copperm}) are surveyed in this
section. Those for question~(\ref{q:copopt}) are discussed in
Section~\ref{sec:surveycopopt} and question~(\ref{q:nocopcert}) is discussed
in Section~\ref{sec:surveycertalgo}.

It may be noted that one way to design an algorithm to test for \COP
is by deriving one from any interval graph recognition algorithm using
the result \cite{hmpv00} \cite[Th~2.7]{d08phd} which demonstrates how
such a derivation can be done.  They use something called {\em Lex-BFS
  ordering} of the vertices of a graph to decide in linear time
whether the graph is an interval graph. They also show how any
interval graph recognition algorithm can be used to recognize matrices
with \COP. This is shown in Theorem~\ref{th:intgraphcop}.

\begin{theoremsansproof}[{\cite[Th.~2]{hmpv00}}]
  \label{th:intgraphcop}
  For a binary matrix $M$ the following statements are equivalent.
  \begin{enumerate}
  \item The row adjacency graph $G_r(M)$ is an interval graph and $M$
    is its maximal clique matrix.
  \item The columns of $M$ are maximal and $M$ has the \COP for rows.
  \end{enumerate}
\end{theoremsansproof}

However, this does not necessarily yield an efficient algorithm. We
will see results that directly solve the problem on matrices since it
is known that questions~(\ref{q:testcop}) and (\ref{q:copperm}) stated
above for \COP are efficiently solvable.  Table~\ref{tab:cophistory}
gives a snapshot of these results.

\tabcophistory

The first polynomial time algorithm for \COP testing was by
\cite{fg65} which uses overlapping properties of columns with \un
s. Their result has close relations to the characterization of
interval graphs by \cite{gh64}. A graph $G$ is an interval graph \iff
all its maximal cliques can be linearly ordered \stt for any vertex
$v$ in $G$, all the cliques that $v$ is incident on are consecutive in
this order. Clearly, this means that the maximal clique incidence
matrix (see
Definition~\ref{def:graphmatrices}~\ref{def::maxcliquematrix}) must
have \COP on rows.

A few years later, a deeply significant result based on very different
ideas in understanding \COP came from Tucker which gave a
combinatorial (negative) characterization of matrices with COP
\cite{at72}. This result influenced most of the \COP results that
followed in the literature including linear time algorithms for \COP
recognition.

\subsection{Tucker's forbidden submatrices for \COP}
\label{sec:tucker}

\cite{at72} discovered certain forbidden structures for convex
bipartite graphs\footnote{The terminology in \cite{at72} differs. It
  uses the term {\em graphs with $V_1$-consecutive arragement} instead
  of {\em convex bipartite graphs}.} and by definition of this graph
class, this translates to a set of forbidden submatrices for matrices
with \cop.  The following are the theorems from \cite{at72} that
acheived this characterization.

Theorem~\ref{th:tuckeratfree} states that convex bipartite graphs
cannot have {\em asteroidal triples} (see
Definition~\ref{def:asteroidal}) contained in the corresponding vertex
partition. This partition of the bipartite graph corresponds to
columns (rows) if its half adjacency matrix has \COP columns (rows).
Theorem~\ref{th:tuckerforbidden} lists the structures in a bipartite
graph that force one of its vertex partitions to have asteriodal
triples -- in other words, it identifies the subgraphs that prevent
the graph from being convex bipartite.

\begin{definition}[Asteroidal triple]
  \label{def:asteroidal}
  % If $G = (V,E)$ is a graph, a set of three vertices from $V$ form
  % an {\em asteroidal triple} if between any two of them there exists
  % a path in $G$ that does not contain any vertex from the closed
  % neighborhood of the third vertex.
  An {\em asteroidal triple} in a graph $G$ is a tuple of vertices
  $\{v_1,v_2,v_3\} \subset V$ such that between every pair of vertices
  from this tuple there exists a path that avoids the closed
  neighbourbood of the third vertex.
\end{definition}

\begin{theoremsansproof}
  [{\cite[Th.~6]{at72}, \cite[Th.~2.3]{d08phd}}]%\\
  A bipartite graph $G = (V_1, V_2, E)$ is convex bipartite on
  columns\footnote{Abridged to match terminology adopted in this
    document. See previous note.} \iff $V_1$ contains no asteroidal
  triple of $G$.
  \label{th:tuckeratfree}
\end{theoremsansproof}

\begin{theoremsansproof}
  [{\cite[Th.~7]{at72}, \cite[Th.~2.4]{d08phd}}]%\\
  In a bipartite graph $G = (V_1, V_2, E)$ the vertex set $V_1$
  contains no asteroidal triple \iff $G$ contains none of the graphs
  $G_{I_k}$, $G_{II_k}$, $G_{III_k}$ (with $k \ge 1$), $G_{IV}$,
  $G_{V}$ as shown in Figure~\ref{fig:forbiddensubgraphs} as subgraphs.
  \label{th:tuckerforbidden}
\end{theoremsansproof}

\figforbiddensubgraphs

Theorem~\ref{th:tuckeratfree} and Theorem~\ref{th:tuckerforbidden}
result in the following Theorem~\ref{th:tuckercop} which characterizes
matrices with \COP.

\begin{theoremsansproof}
  [{\cite[Th.~9]{at72}, \cite[Th.~2.5]{d08phd}}]%\\
  A matrix $M$ has \COP \iff it contains none of the matrices 
$M_{I_k}$, $M_{II_k}$, $M_{III_k}$ (with $k \ge 1$), $M_{IV}$,
  $M_{V}$ as shown in Figure~\ref{fig:forbiddensubmatrices} as submatrices.
  \label{th:tuckercop}
\end{theoremsansproof}

\figforbiddensubmatrices

It can be verified that the matrices in
Figure~\ref{fig:forbiddensubmatrices} are the half adjacency matrices
of the graphs in Figure~\ref{fig:forbiddensubgraphs} respectively
which is not surprising due to
Definition~\ref{def:graphwithcop}~\ref{def::convexbi}.

\subsection{Booth and Lueker's $PQ$-tree -- a linear COT algorithm} %[\PQtree]

Booth and Lueker in their paper \cite{bl76} gave the first linear
algorithm for \cop testing while given a linear time interval graph
recognition algorithm by a simplification of \cite{lec67}'s planarity
test algorithm. This \COT algorithm has time complexity
$O\left(m+n+f\right)$ where $m \times n$ is the order of the input
matrix and $f$ is the number of $\un$s in it.
% Hsu in wlh01 said BL76 gave a planarity test - "the induced PQ-tree
% algorithm can considerably simplify Booth and Lueker’s modification
% of Lempel, Even and Cederbaum’s planarity test. " 
%
% McConnell in hm03 said the first linear algo for interval graph
% recog was by BL76 - "Benzer’s work and other applications of
% interval graphs motivated a search for eﬃcient algorithms for
% recognizing interval graphs, and for constructing a set of intervals
% to represent the graph when it is one [8]. A linear-time algorithm
% was given in 1976 by Booth and Lueker [2]." 
\cite{bl76} introduces a data structure called \PQtree and their \COP
testing algorithm is a constructive one that outputs a \PQtree if the
input has \COP. A \PQtree represents all the \COP orderings of the
matrix it is associated with. \cite{bl76}'s algorithm uses the fact
that if a matrix has \COP, a \PQtree for it can be constructed. It is
interesting to note that aside from interval graph recognition and
\COP testing, \PQtree is also useful in other applications like
finding planar embeddings of planar graphs \cite{lec67,mcc04} and
recognizing \CROP in a matrix.

%\begin{minipage}[h]{5.5in}
\begin{definition}[\PQtree \cite{bl76, mcc04}]
  A \PQtree of matrix $M$ with \COP on columns (rows), is a tree with
  the following properties.
  \begin{enumerate}[i.]
    \singlespacing
  \item Each leaf uniquely represents a row (column) of $M$. The leaf
    order of the tree gives a \COP order for column (row) (\COP order
    for column requires permutation of rows and vice versa.) for $M$.
  \item Every non-leaf node in the tree is labeled $P$ or $Q$.
  \item \label{def::nodep} The children of $P$ nodes are
    unordered. They can be permuted in any fashion to obtain a new
    \COP order for $M$.
  \item \label{def::nodeq} The children of $Q$ nodes are linearly
    ordered. Their order can be reversed to obtain a new \COP
    order for $M$.
  \end{enumerate}
  \label{def:pqtree}
\end{definition}
%\end{minipage}

See Figure~\ref{fig:pqtree} for an example of \PQtree. It may be noted
that there is no way an empty set of \COP orderings can be represented
in this data structure. For this reason, \PQtree is undefined for
matrices that do not have \COP.  Thus effectively, there exists a
bijection between set of matrices with \COP and the set of \PQtrees
(accurately speaking, each matrix with \COP bijectively maps to an
equivalence class of \PQtrees resulting from properties (\ref{def::nodep}) and
(\ref{def::nodeq})).

\figpqtree

The \cite{bl76} algorithm with input $n \times m$ matrix $M$ starts
with a \PQtree for a vacuous $n \times 0$ matrix $M'$ (submatrix
induced by 0 columns). This is known as a {\em universal} \PQtree
which is one with its root as a $P$ node and only leaves as its
children -- each leaf representative of a row of input (by definition
of \COP for columns). This induced submatrix $M'$ vacuously has \COP.
Each column is then added iteratively to $M'$ to check if the new $M'$
has \COP.  By a complicated, but linear, procedure the algorithm does
one of the following actions in each iteration: (a) declare that $M$
has no \COP, or (b) modify the current \PQtree to represent the new
$M'$ (which clearly, must have \COP, since if not, option (a) would
have been executed).

Judging from notes in literature, this algorithm is apparently
notoriously difficult to program. In the procedure to modify the
\PQtree at each iteration, nodes are considered from leaves to
tree. At each node considered, it uses one of nine templates to
determine how the tree must be altered in the vicinity of this
node. Recognition of this template poses a difficult challenge in
terms of implementing it. Each template is actually a representative
of a larger class of similar templates, which must be dealt with
explicitly by a program\cite{mcc04}.

After the invention of \PQtrees, presumably due to the implementation
challenge it posed, there has been several variants of the same in the
literature, like \PCtree \cite{sh99,wlh01,hm03}, generalized \PQtree
\cite{km89,mcc04}, \PQRtree \cite{mm96,mpt98} etc. Most of these are
generalizations of \PQtree\ -- for instance, \PCtree is generalized to
matrices with \CROP, \PQRtree and generalized \PQtree are generalized
to matrices and set systems with or without \COP. \cite{km89} invented
a modified form of \PQtree, called $MPQ$-trees, a simpler incremental
update of the tree only for recognizing interval graphs.  \cite{kr88}
constructed efficient parallel algorithms for manipulating \PQtrees.

In the next few sections we will see some of these variations.

\subsection{\PQRtree\ --  COT for set systems}
\label{sec:surveycertalgo}

Section~\ref{sec:motive} mentions how a binary matrix naturally maps
to a system of sets.  A set can be constructed for each column of
matrix with its elements being those row indices at which the column
has \un s. Thus the collection of sets corresponding each column of
the matrix forms a set system with universe as the set of all row
indices of the matrix.  This simple construction is formally described
in Definition~\ref{def:matrixsetsystem} along with the idea of \cop
for set systems (As seen in Section~\ref{sec:courseschedule}).

\begin{definition}[{\em \Cop for set systems}]%
  \label{def:matrixsetsystem}%
  % [\emph{Set system of a binary matrix}.]
  Let $M$ be a binary matrix of order $n \times m$ and $\{c_i \mid i
  \in [m]\}$ be the columns in $M$.  A set system $\cF_{M} = \{S_i
  \mid S_i \subseteq [n], i \in [m]\}$ is defined such that for every
  column $c_i$ of $M$, set $S_i = \{j \mid m_{ji} = \un \}$. The
  collection $\cF_M$ is the {\em set system of binary matrix} $M$. The
  {\em binary matrix for set system} $\cF$ is conversely constructed
  and denoted by $M^\cF$. Thus, $M^{\cF_M} = M$.%
  \par\noindent%
  A set system $\cF$ from universe $U$, $|U| = n$ has the {\em \cop}
  if there exists a linear order or permutation $\sigma = w_1w_2\ldots
  w_n$ that can be applied to $U$ \stt each set $S \in \cF$ becomes a
  consecutive subsequence (also termed an {\em interval})
  $w_{i}w_{i+1}\ldots w_{i+k-1}$ on $\sigma$ for some positive integer
  $i \le n+1-k$ where $k = |S|$.
  \par\noindent%
  The set $\valid{U,\cF}$ respresents all \COP orders of $\cF$ in $U$.
  \dstop
\end{definition}

It is easy to see the equivalence of this definition to \COP for
matrices in Definition~\ref{def:copmatrix}.

Before proceeding to describe \PQRtree per se, we will see a few more
terminologies that will make the subsequent discussion in this section
simpler.

\begin{definition}[\em Orthogonal sets \cite{n89, mm96,
    mcc04}\footnote{\cite{mcc04} does not use the term ``mutually
    orthogonal'' but refers to the same idea as ``sets that do not
    overlap''. This terminology is also used in other literature like
    \cite{nsnrs09,wlh02}.} etc.]
  \label{def:orthosets}%
  Let $\cF$ be a set system with universe $U$ and sets $A, B \in \cF$.
  \begin{enumerate}
  \item $A$ and $B$ are said to have a {\em trivial intersection} if
    $A \cap B$ is $\emptyset$ or $A$ or $B$.  In other words, $A$ and
    $B$ are either disjoint or one is the subset of the other.

  \item $A$ and $B$ are called {\em mutually orthogonal} or $A$ is {\em
      orthogonal to} $B$ and vice versa, if they have a trivial
    intersection.

  \item {\em Trivial subsets} of a universe $U$, denoted by $\cT(U)$,
    are sets that have trivial intersections with any set in
    $\power{U}$. These sets are $U$, singleton sets in
    $\power{U}$ and $\emptyset$ \cite{n89, mm96}.
    Thus, $\cT(U) = \{U\} \  \bigcup \  \{ \{v\} \mid v \in U \} \  \bigcup \
    \{\emptyset\}$.

  \item {\em Orthogonal sets of a set
      system}\footnote{\cite[Def.~3.1]{mcc04} uses the term {\em
        non-overlapping family} of $\cF$ and denotes it by
      $\cN(\cF)$. \label{mcc1}} $\cF$ with universe $U$ are subsets of
    $U$ that are orthogonal to all sets in $\cF$. The set of all
    orthogonal sets to $\cF$ is denoted by $\cF^{\bot}$.

  \item $\cF$ is called {\em complete}\footnote{\cite{mcc04} calls
      this a {\em weakly partative family}. \label{mcc2}} if the
    following hold true for every pair of non-orthogonal (overlapping)
    sets $A, B$ in $\cF$.

    \begin{tabular}[h]{lllll}
%    \begin{enumerate}[i.]
    (i) $\cT(U) \subset \cF$ &
    (ii) $A \cup B \in \cF$ &
    (iii) $A \cap B \in \cF$ &
    (iv) $A \setminus B \in \cF$ &
    (v) $B \setminus A \in \cF$
%   \end{enumerate}      
    \end{tabular}

    In other words, $\cF$ contains all the trivial
    subsets of $U$, $A \cup B$ and the partitions of $A \cup B$
    defined by intersection and set difference.

  \item $\overline{\cF}$ represents the smallest super set system of
    $\cF$ that is complete\footnote{\cite[Def.~3.2]{mcc04} calls
      this the {\em weak closure} of $\cF$ denoted by
      $\cW(\cF)$. \label{mcc3}}

  \item \label{def::slashop} The binary operator $/$ is defined as
    follows. Let $A, B$ be two sets from set sysetm $\cF$.
    \begin{align*}
      A/B &= A \setminus B \cup \{b\}\\
      \cF/B &= \{ S/B \mid S \in \cF, S \nsubseteq B \}
    \end{align*}
    $b$ is a new element added as a representative for $B$.
 \end{enumerate}
  \dstop
\end{definition}

An important observation made by \cite{mm96} is presented now along
with a few theorems that help in decomposing the \COP problem on $\cF$
into subproblems.

\begin{observation}[{\cite[Sec.~3]{mm96}}]
  If $\cF$ is a set system with \COP then, after applying the \COP
  order, not only must every set in $\cF$ be consecutive but the
  following sets must also be consecutive for any $A, B \in \cF$.
\begin{enumerate}
  \item The intersection $A \cap B$ 
  \item The union $A \cup B$ if $A \cap B \ne \emptyset$
  \item The relative complements $A \setminus B$ and $B \setminus A$
    if $B \nsubseteq A$ and $A \nsubseteq B$ respectively.
  \item Also note that trivially, sets in $\cT(U)$ are consecutive in
    any permutation of $U$\footnote{$\emptyset$ is considered
      consecutive by convention.}.
\end{enumerate}
\end{observation}

The following theorem gives a very interesting property of orthogonal
sets and $\overline{\cF}$

\begin{theoremsansproof}[{\cite[Th.~3,6]{mm96}}]
  \label{th:validcop}
  For any set system $\cF$ we have the following. \par
  \centering
    $\valid{\cF} = \valid{\overline{\cF}}$ \\
    $\cF^{\bot} = \overline{\cF^{\bot}} = (\overline{\cF})^{\bot}$
\end{theoremsansproof}

\begin{proposition}
  \label{pr:validcop}
  Any set that is consecutive on all the \COP permutations of $\cF$ is
  present in $\overline{\cF}$.
\end{proposition}

Proposition~\ref{pr:validcop} is owing to the following theorem by
which \cite{mm96} describes a way to decompose the problem of finding
all \COP orders of $\cF$ into two subproblems using sets in
$\overline{\cF} \cap \cF^{\bot}$. The operator $/$ used is defined in
Definition~\ref{def:orthosets}~(\ref{def::slashop}).

\begin{theoremsansproof}[{\cite[Th.~7]{mm96}}]
  \label{th:mmdecomp}
  For any set system $\cF$, and $\emptyset \ne H \in \overline{\cF}
  \cap \cF^{\bot}$ we have the following.\par
  \centering
  $\valid{U,\cF} = \valid{U/H,\cF/H} * \valid{H,\cF \cap 2^H}$
\end{theoremsansproof}

\def \ff {\overline{\cF} \cap \cF^{\bot}}
The idea behind Theorem~\ref{th:mmdecomp} is as follows. A permutation $\alpha$
of $U$ is a composition of two permutations with respect to $H$ - (i) a
permutation $\gamma$ of $H$ and (ii) a permutation $\beta$ of $U
\setminus H$.

For two mutually orthogonal sets $A, B$ such that $A \nsubseteq B$,
$A/B$ is defined as the set obtained by removing all elements of $B$
from $A$ and adding a representative element for $B$ in $A$. Being orthogonal,
$A/B$ results in only the following three possibilities.
\begin{enumerate}[i. ]
\item $A$ and $ B$ are disjoint: $A/B  = A$
\item $A$ is a subset of $B$: $A/B$ is not defined
\item $B$ is a subset of $A$: $A/B = A \setminus B \cup \{b\}$, where $b$
is a new element not in $U$ added to represent $B$.
\end{enumerate}

We observe that this idea of decomposing the \COP problem into two
subproblems in \cite{mm96} is very similar to the substitution
decomposition of a set system given in \cite[Sec.~4]{mcc04}.\footnote{Using
  the theory cited in footnotes~\ref{mcc1},\ref{mcc2},\ref{mcc3}.}

The following corollary, Corollary~\ref{cor:pqrrecursive} states how
\PQRtree elegantly fits into this whole theory and help in computing
all \COP orders of $\cF$.

\begin{corollary}[{\cite[Cor.~8]{mm96}}]
  \label{cor:pqrrecursive}
  Let $\cF$ be a set system with universe $U$ and $H$ is a non-empty
  orthogonal set $H \in \ff$. If there is a \PQRtree $T_1$ that
  encodes all permutations in $\valid{U/H, \cF/H}$ and also a \PQRtree
  $T_2$ that encodes all permutations $\valid{H,\cF \cap \power{H}}$,
  then a \PQRtree $T$ for $\cF$ can be obtained by replacing the leaf
  $h$ in $T_1$ by $T_2$, where $h$ is the representative element for $H$.
\end{corollary}

%Finding elements in $\ff$ is done using \PQRtree.

Thus we have a recursive algorithm that can compute the \PQRtree for
$\cF$ provided we find an element from $\ff$ in each iteration. The
non-empty sets in $\ff$ are called {\em node sets} since they form the
nodes in the \PQRtree, as alluded to by
Corollary~\ref{cor:pqrrecursive}. They are calculated as follows.
First method is by computing the overlap components of $\cF$. The
overlap components is the partition that results from the {\em
  overlap} equivalence relation which is nothing but {\em
  non-orthogonal} equivalence relation (It is easy to verify
  that this relation is indeed an equivalence relation.). Overlap
components are linearly computable\cite{mm95,wlh92}. Once these
elements are factored out, the rest of the node sets are obtained by
the second method which is identifying {\em twin}
elements\footnote{\cite[Sec.~3]{kklv10} calls this {\em indistinguishable}
  elements. The equivalence class is called a {\em slot}.}. Two
elements $a, b \in U$ are twins if their membership in every set of
$\cF$ is in tandem with each other, \ie $\{a,b\} \bot \cF$. This is
clearly an equivalence relation and their equivalence classes is known
to be computable in linear time\cite{wlh01} and even in
logspace\cite{kklv10}.

The recursion end condition is when one cannot find any more sets from
$\ff$ that are non-trivial. This is when $\ff = \cT(U)$. This is the
point where the parents of the leaves of the final \PQRtree are created. The
following theorem helps the algorithm decide whether a $P$, $Q$ or
$R$ node must be created.

\begin{theoremsansproof}[{\cite[Th.~9]{mm96}}] %,\cite[Th.~2.1]{mcc04}
\label{th:pqrfinalmm}
If $\cF$ is a set system with universe $U$ and $|U| \ge 3$ then one of
the following statements hold.
\begin{enumerate}
\item \label{th::it1} $\overline{\cF} = \cT(U)$ 
\item \label{th::it2}$\overline{\cF} = consec(\alpha)$ for
  some permutation $\alpha$ on $U$ ($consect(\alpha)$ is a linear order)
\item \label{th::it3}$\overline{\cF} = \power{U}$ 
\end{enumerate}
\end{theoremsansproof}

In case (\ref{th::it1}) of Theorem~\ref{th:pqrfinalmm} all elements in
$U$ are made children of a $P$ node. In case (\ref{th::it2}) all
elements in $U$ are made children of a $Q$ node in the order given by
$\alpha$. Finally, case (\ref{th::it3}) is the situation where no permutation
of $U$ gives \COP. All elements of $U$ are in this case made children
in an $R$ node.

\subsubsection{Comparision with generalized \PQtree}

Stepping back a little, the reader may have noticed the cross
references between \PQRtree and other data structures in footnotes
earlier.  Generalized \PQtree or \gPQtree is a data structure defined
in \cite{n89} to represent all orthogonal sets of a set system
$\cF$. A data structure with the same name was later defined in
\cite{mcc04} as part of a {\em substitution decomposition} for a set
system $\cF$ and subsequently \cite{mcc04} gives a new
characterization of $\cF$ using a so-called incompatibility graph
(this is discussed in more detail in
Section~\ref{sec:incompatibilitygr}). As seen above, \PQRtree is
defined in \cite{mm96} as a data structure to represent any set system
$\cF$ with additional information in their $R$-nodes if $\cF$ has no
\COP. All three of these data structures are proposed as
generalizations of \cite{bl76}'s \PQtree and hence produce the \PQtree
if $\cF$ has \COP. As a whole, all these three data strutures are
largely identical with their differences being notional. We discussed
the basic theory that they all hold and predominantly used the
terminology from \cite{mm96}. To the best of our knowledge this
comparison study has not been made earlier in the literature.  We
observe the three data strutures of \PQRtree, \gPQtree and generalized
\PQtree to be equivalent. The terminologies of the theory of all these
data structures is summarised in Table~\ref{tab:pqrcomparison}.

Going back to the theory we were discussing earlier,
Theorem~\ref{th:pqrfinalmm} and the theory leading to it is very
similar to \cite[esp. Th.~2.1, 3.5. also Th.~3.2, 3.3, 3.4]{mcc04}
which categorizes the nodes in the above the three cases as {\em
  prime}, {\em linear} and {\em degenerate}
respectively. \cite{mcc04}'s generalized \PQtree is created in similar
ways as \PQRtree above. This tree is, in essence, the Hasse diagram of
what they call {\em strong elements} of weak closure, $\cW(\cF)$(\ie
$\overline{\cF}$). Strong elements of a set family are elements that
do not overlap with any other elements in the family, \ie it is
orthogonal to all other sets\cite[Def.~3.3]{mcc04}.

\begin{theoremsansproof}[{\cite{mm96}, \cite[Th.~3.6]{mcc04}}]
The set system $\cF$ has \COP \iff its \PQRtree has no $R$ nodes.  
\end{theoremsansproof}

Thus \PQRtree gives a data structure that encodes possible linear
orderings of a set that demonstrates the \COP property in it or
narrows it down to parts of the universe in $R$ nodes that prevent the
set system from having \COP.

We will now see, in brief, one more generalization of \PQtree before
seeing another approaches to solving \COP testing called ICPIA which
is a set theoretic characterization of \COP.

\begin{table}[t]
  \def\colwidth{4cm}
  \centering
  \fontencoding{OT1}\fontfamily{cmss}\selectfont%
  \footnotesize%
  \setlength\extrarowheight{0.1in}

  \begin{tabular}{l >{\columncolor{\tblhcolor}}l l}

    \rowcolor[gray]{0.8} %                 Header row colored
    \normalsize \gPQtree \cite{n89} & 
    \normalsize \PQRtree \cite{mm96} & 
    \normalsize generalized \PQtree \cite{mcc04}
    \\

    Trivial intersections &
    Trivial sets $\cT(U)$ & 
    \\\hline

    Trivially intersecting sets &
    Mutually orthogonal sets, $A \bot B$ &
    Non-overlapping sets
    \\\hline

      &
    Complete collection &
    Weakly partative family
    \\\hline

    &
    $\overline{\cF}$&
    Weak closure, $\cW(\cF)$
    \\\hline

    &
    $\cF^{\bot}$&
    Non-overlapping family, $\cN(\cF)$
    \\\hline

    &
    \parbox[t]{\colwidth}
    { 
      Sets in $\cF$ orthogonal to all other sets in $\cF$ 
    }&
    Strong elements
    \\\hline

    &
%     \parbox[t]{\colwidth}
%       {}
    \PQRtree
    &
    \parbox[t]{\colwidth}
    {
      Decomposition tree of $\cW(\cF)$, $T(\cW(\cF))$
    }
    \\\hline

    &
    Node sets, $(\cF \cap \cF^\bot) \setminus \{\emptyset\}$
    \\\hline

    &
    \parbox[t]{\colwidth}
    {
      $P$-node $\Leftrightarrow \cF = \cT(\cF)$\\
      $Q$-node $\Leftrightarrow \cF = consec(\alpha)$\\
      $R$-node $\Leftrightarrow \cF = \power{U}$\\
    }
    &
    \parbox[t]{\colwidth}
    {
      {\em Prime}: For every $I$, $\emptyset \ne I \subset [k],
      \cup_{i \in I}S_i \notin \cF$\\
      {\em Linear}: There exists an ordering of $[k]$
      s.t. $\emptyset \ne I \subset [k]$ implies $\cup_{i \in I}S_i
      \in \cF$ only if members of $I$ are consecutive in the ordering\\
      {\em Degenerate}: For every $I$, $\emptyset \ne I \subset [k],
      \cup_{i \in I}S_i \in \cF$\\
    }

    \\\lasthline

  \end{tabular}
  \caption[\figtabsize Comparison of theory of \PQRtree, \gPQtree, generalized
  \PQtree]{\figtabsize Comparison of theory of \cite{mm96} \PQRtree, \cite{n89} \gPQtree,
    \cite{mcc04} generalized \PQtree}
  \label{tab:pqrcomparison}
\end{table}

\subsection{\PCtree -- a generalization of \PQtree}

\PCtree is another generalization of \PQtree. It is a data structure that is
analogous to \PQtree but for matrices with \crop. \PCtree was
introduced by \cite{sh99} for the purpose of planarity testing where
this data structure represents partial embeddings of planar
graphs. % noted from wlh01 abstract.
In \cite{wlh01}, Hsu reintroduces \PCtree as a generalization of
\PQtree and shows how it simplifies \cite{bl76}'s planarity test by
making the \PQtree construction much less complicated. Later
\cite{hm03}, discovers that \PCtree is a representation of all \crop
orders of a matrix when it is unrooted. \PCtree presented in
\cite{wlh01} is rooted; however the construction of \PCtree is the
same in both results. The property of being unrooted is necessary in
order to use \PCtree as a data structure for encoding circular
ordering. Definition~\ref{def:pctree} defines \PCtree.

\begin{definition}[\PCtree \cite{wlh01,d08phd}]
  \label{def:pctree}
   A \PCtree of matrix $M$ with
  \CROP on columns (rows), is a tree with the following properties.
  \begin{enumerate}[i.]
    \singlespacing
  \item Is unrooted -- thus it has (a) no parent child relationship
    between nodes (b) there is no left to right (or vice versa)
    ordering.
  \item Each leaf uniquely represents a row (column) of $M$. The leaf
    order of the tree gives a \CROP order for column (row) for
    $M$. Moreover, any sequence obtained by considering the leaves in
    clockwise or counter-clockwise order describes a \CROP order for
    $M$.
  \item Every non-leaf node in the tree is labeled $P$ or $C$.
  \item The neighbors of $P$ nodes can be permuted in any fashion to
    obtain a new \CROP order for $M$.
  \item \label{def::pcmirror} The tree can be changed by applying the following
    ``mirroring'' operation to obtain a new \CROP order for $M$. Root
    the \PCtree at a neighbor of a $C$-node, $v$ and mirror the
    subtree whose root is $v$ and finally unrooting the
    tree. Mirroring a subtree is done by putting the children of every
    node of the subtree in reverse order.
  \end{enumerate}
\end{definition}

\begin{figure}[htbp]
  \centering
  %\includegraphics[scale=0.4]{../img/pctree_hm03.png}

  \begin{tabular}[h]{lcc}
    \pctreegraphmatrix & 
    \pctreegraphi &
    \pctreegraphii
  \end{tabular}
  \caption[\figtabsize \PCtree]{\figtabsize \PCtree with the
    associated matrix. The dark circular nodes are $P$-nodes and the
    double circular nodes are $C$-nodes.  The two \PCtrees are both
    for the same matrix. The second tree is obtained by performing the
    ``mirroring'' operation on the subtree (highlighted in the figure)
    rooted at the bottom right $C$-node.  \cite[Fig.~2.11]{d08phd}}
  \label{fig:pctree}
\end{figure}

As a data structure when \PCtree is compared with \PQtree, the
differences are, (i) it is unrooted, (ii) it represents all \CROP order of a
matrix (iii) it has $C$ nodes instead of $Q$ nodes which can be
``mirrored'' (operation defined in Definition~\ref{def:pctree}~\ref{def::pcmirror}).
The algorithms of construction of \PQtree in \cite{bl76} and that of
\PCtree in \cite{wlh01, hm03} starkly differ since the latter is a
much simplified procedure.

% WLH01 abstract:
% [..] The original implementation of the PQ-tree algorithms by Booth and
% Lueker using nine templates in each bottom-up iteration is rather
% complicated. Also the complexity analysis is rather intricate. We give
% a very simple linear time PC-tree algorithm with the following
% advantages: (1) it does not use any template; (2) at each iteration,
% it does all necessary tree-modification operations in one batch and
% does not involve the node-by-node bottom-up matching; (3) it can be
% used naturally to test the circular ones property in matrices; (4) the
% induced PQ-tree algorithm can considerably simplify Booth and Lueker’s
% modification of Lempel, Even and Cederbaum’s planarity test.

% \tnote[FYI]{ The results in cite~{bl76} on COT are based on the
%    result that interval graphs are AT-free chordal graphs cite~{lb62}?
%  No they don't seem related. this is the stupid bl lb confusion.}%

\subsection{ICPIA - a set cardinality based \COP test}
\label{sec:icpiasurvey}
In this section we see a characterization that does not use
\PQtree. \cite{nsnrs09} describes a characterization of \cop solely
based on the cardinality properties of the sets in the set system and
does not use any variants of \PQtrees.

\begin{definition}[Intersection Cardinality Preserving Interval
  Assignment (ICPIA)]
  \label{def:icpia}
  Let $\cF = \{A_i \mid A_i \subseteq U, i \in [m]\}$ be a set system
  from universe $U$ and the set of ordered pairs $\Pi = \{(A_i, B_i)
  \mid B_i  \text{ is an interval from } [n], i \in [m]\}$ be an {\em interval
    assignment} of $\cF$, then $\Pi$ is called an {\em ICPIA} if it has
  the following properties.
  \begin{enumerate}[i.]
  \item $|A_i| = |B_i|$ for all $i \in [m]$
  \item $|A_i \cap A_j| = |B_i \cap B_j|$ for all $i, j \in [m]$
  \end{enumerate}
\end{definition}

\begin{theoremsansproof}[{\cite[Th.~1]{nsnrs09}}]
  \label{th:icpia1}
  If an interval assignment $\Pi$ is feasible, then it is an ICPIA.
\end{theoremsansproof}

The necessity condition of Theorem~\ref{th:icpia1} is fairly
obvious. It turns out that it is sufficient too.  This is demostrated
by two algorithms which work as follows. The first one iterates over
the assignment $\Pi$ %set system $\cF$
and changes it in each iteration till the altered
assignment %set system
has no more overlapping intervals (intervals are the second
  element in each ordered pair of $\Pi$. Def.~\ref{def:icpia}) %sets
\cite[Alg.~1]{nsnrs09}. This is acheived by replacing two overlapping
assignment pairs $(A_i, B_i)$ and $(A_j, B_j)$ with the partitions of
$(A_i \cup A_j,B_i \cup B_j)$ induced by their overlaps -- $(A_i
\setminus A_j, B_i \setminus B_j)$, $(A_i \cap A_j, B_i \cap B_j)$ and
$(A_j \setminus A_i, B_j \setminus B_i)$.

To see it in terms of the theory explained in
Section~\ref{sec:surveycertalgo}, this results in the weak closure of
$\cF$. Let $\cF$, $\cI$ be the set system and interval
system (A set system where all the sets are intervals.),
respectively, involved in the assignment $\Pi$ and $\cF'$, $\cI'$ be
the same for the output of the algorithm $\Pi'$.  It can be observed
that $\cF' \cup \cF$ is $\overline{\cF}$ or the weak closure of
$\cF$. The analog is true for $\cI \cup \cI'$.  Moreover, this output
$\cF'$, $\cI'$ contains the strong elements of $\overline{\cF}$,
$\overline{\cI}$ respectively. It turns out that $\cI'$ is also an
interval system and that $\Pi'$ is an ICPIA \cite[Lem.~2]{nsnrs09}.

The next algorithm \cite[Alg.~2]{nsnrs09} further refines $\Pi'$ to
$\Pi''$ which represents the family of permutations that yield the
\COP orders represented by $\Pi$ of $\cF$. The subset relation Hasse
diagram of $I'$ is created with each node representing the
corresponding assignment pair of the interval from $\Pi'$. Since all intervals in
$\Pi'$ are non-overlapping, this is a tree. Notice that this means any
two intervals are now either disjoint or one is contained in the
other. Hence this tree is called a {\em containment tree}.  The
algorithm traverses this tree in post order fashion where the post
order function is to replace an assignment pair $(X,Y)$  in $\Pi'$ by their child
assignment pairs in the containment tree, say $(X_1, Y_1),(X_2, Y_2)
\ldots (X_k,Y_k)$. The result $\Pi''$ gives the
\COP order for $\Pi$. These two algorithms prove that ICPIA is a
sufficient condition for the feasibility of the interval assignment.

\begin{theoremsansproof}[{\cite[Th.~2]{nsnrs09}}]
  \label{th:icpia2}
  Let  $\cF = \{A_i \mid A_i \subseteq U, i \in [m]\}$ be a set system
  from universe $U$, $|U| = n$
  and $\Pi = \{(A_i,B_i) \mid B_i \text{ is an interval from } [n], i
  \in [m]\}$ be an ICPIA. Then there exists a permutation $\sigma: [n]
  \rightarrow [n]$ \stt $\sigma(A_i) = B_i$.
\end{theoremsansproof}

What we saw so far in this section is checking for the feasibility of
a given interval assignment. ICPIA can also be used to find an
interval assignment when only the set system $\cF$ is given. Part of
the approach is similar to the overlap component idea in \cite{fg65,
  wlh02}. Once the overlap components are factored out, they can be
independently assigned a subinterval to which the sets in the
component need to be mapped. This is simple because the overlap
components are disjoint from each other. Now the subproblem is finding
\COP order for each overlap component. In each component, the set
whose overlaps with the rest of the sets in that component forms a
single inclusion chain (see Definition~\ref{def:poset}. In this
particular reference, the set of sets is the set of overlaps.)  is
chosen first and assigned the leftmost\footnote{This is by
  convention. Rightmost can also be chosen analogously to get a
  different \COP order.} interval. The next candidate set is chosen
from this set's overlapping sets. The interval assigned is calculated
by simple intersection cardinality means. Assume that the next set
overlaps on one of the ``ends'' of the current set and calculate the
interval by shifting left or right of the current interval and
adjusting its size to match the next set's cardinality. Once the next
set is assigned an interval, the current assigned sets are all checked
for ICPIA. This is a backtracking algorithm -- if at any point ICPIA
fails, another overlapping set is chosen. If all overlapping set
attempts fail ICPIA, a different ``first set'' is chosen. This is a
remarkably simple algorithm due to its obvious simplicity in
implementation.

\section{Matrices without \COP}
\label{sec:surveycopopt}

\COP is a very beneficial property since it simplifies the structure
of the input (binary matrix or set system) leading to
efficient algorithms to otherwise hard problems.  While
Section~\ref{sec:surveycoptest} discusses matrices with \COP and how
to compute \COP orders, this section describes a couple of problems
when the input does not have \COP and identify their forbidden
substructures that prevents them from having \COP.
%describes the problems  that are
% of interest when a matrix does not have this property

\subsection{Finding forbidden submatrices}
\label{sec:forbiddensubalgo}

It is known from Theorem~\ref{th:tuckercop} that when a matrix does
not have \COP, it must have one of Tucker's forbidden submatrices (See
Section~\ref{sec:tucker} for detailed discussion on Tucker's forbidden
substructures for matrices with \COP). However, none of the \COP test
algorithms discussed in Section~\ref{sec:surveycoptest} outputs the
forbidden submatrix when the test is negative. This perhaps gives a
hint about how hard it is to find these substructures in the input
matrix. Using a reduction to the longest path problem, \cite{v85}
proves that finding a Tucker's forbidden submatrix of at least $k$
rows is NP complete. An example of an optimization problem for a
matrix without \COP is finding if deleting at most $k$ rows from input
will restore \COP. This is known to be NP complete
\cite{b75-phd}. There are several such optimization problems that try
to find ways to get \COP on the input matrix with certain alterations
and finding smallest forbidden submatrix is at the heart of these problems.

In this section we will summarize \cite{d08phd}'s approximation
algorithm for finding forbidden submatrix. We will use $\Tuckerset$ to
refer to the set of Tucker's forbidden submatrices shown in
Figure~\ref{fig:forbiddensubmatrices}.

A corollary to one of Tucker's theorems presented in
Theorem~\ref{th:tuckeratfree} is given in
Corollary~\ref{cor:tuckertheorem}. This is central to \cite{d08phd}'s
algorithm.

\begin{corollary}[{\cite[Cor.~3.1]{d08phd}}]
  \label{cor:tuckertheorem}
  A matrix $M$ has \COP \iff its representing bipartite graph $G_M$
  (the graph whose half adjacency matrix is $M$) does not contain an
  asteroidal triple whose three vertices correspond to columns of M.
\end{corollary}

Using Corollary~\ref{cor:tuckertheorem}, a forbidden submatrix can be
found as follows. For every vertex triple $x, y, z$ in $G_M$ that
correspond to columns of $M$, determine the sum of the lengths of
three shortest paths connecting each pair of these vertices (there are
3 possible pairs) avoiding the neighborhood of third vertex. If all
three paths exist, then $x, y, z$ indeed form an asteroidal
triple. Among all such triples, select the triple $u, v, w$ where the
sum is minimum and return the rows and columns of $M$ that correspond
to the vertices in the three shortest paths described earlier for this triple. This
return value forms a submatrix and it obviously must contain a
submatrix from $\Tuckerset$ because it represents an asteroidal
triple. It may be noted that this procedure does not always return a
submatrix of smallest size because the shortest paths computed may not
be vertex disjoint. The following claim shows that this certainly
gives a reasonable approximation of the solution. The claim has been
abridged to summarize the main idea behind this algorithm.
Note that a $(x,y)$-matrix is one with at most $x$ number of {\un}s in every
column and at most $y$ number of {\un}s in every row. If $x$ or $y$ is given as
$*$, it denotes that number of {\un}s is unbounded.

\begin{proposition}[{\cite[Pr.~3.1, {\em abridged}]{d08phd}}] 
  \label{prop:tuckerdomapprox}
  Let $M$ be a $(*,\Delta)$-matrix of size $m \times n$ that contains
  a forbidden submatrix $M'$ of size $m' \times n'$ from $\Tuckerset$. Then
  this algorithm computes a submatrix that belongs to $\Tuckerset$
  with at most $m'+5$ rows and $n'+3$ columns in $O(\Delta m n^2 +
  n^3)$ time.
\end{proposition}

In fact, the approximation mentioned in
Proposition~\ref{prop:tuckerdomapprox} is the worst case when
the actual solution $M'$ is of type $M_{IV}$. The approximation is
slightly better for the other types of submatrices from $\Tuckerset$,
the best case being an exact solution when $M'$ is $M_{I_k}$ or
$M_{II_k}$. Thus this algorithm works accurately when the solution is of
type $M_{I_k}$ or $M_{II_k}$.

Next, notice that Tucker matrix types $M_{IV}$ and $M_{V}$ are
constant size -- $4 \times 6$ and $4 \times 5$ respectively. Thus they
can be searched for in polynomial time even with brute force method as
Corollary~\ref{cor:mivmvpolytime} states.

\begin{corollary}[{\cite[Cor.~3.3, {\em abridged}]{d08phd}}] 
  \label{cor:mivmvpolytime}
  Let $M$ be a binary $(*, \Delta)$-matrix of size $m \times n$ and
  $M'$ be a binary matrix from $\Tuckerset$. A submatrix in $M$ that
  is isomorphic to $M'$ can be found in $O(\Delta^3m^2n^3)$ time or
  $O(\Delta^4m^2n)$ time when $M'$ is of type $M_{IV}$ or $M_V$
  respectively.
\end{corollary}

Their algorithm for finding $M_{III_k}$ type submatrix is slightly
involved and we only present here the analysis result in
Proposition~\ref{cor:miiitime}. The algorithm is derived from the
structural similarities between the first three types of Tucker
matrices and the fact that $M_{I_k}$ is a chordless cycle of length
$2k+4$.

\begin{proposition}[{\cite[Pr.~3.6]{d08phd}}] 
  \label{cor:miiitime}
  Let $M$ be a $(*,\Delta)$-matrix $M$ of size $m \times n$. Then a
  minimum size submatrix of type $M_{III_k}$ in $M$ can be found in
  $O(\Delta^3m^3n+\Delta^2m^2n^2)$ time.
\end{proposition}

All the above algorithmic results are combined to obtain the final
single algorithm for finding forbidden
submatrix. Theorem~\ref{th:forbiddenalg} gives this result.

\begin{theoremsansproof}[{\cite[Th.~3.1]{d08phd}}]
  \label{th:forbiddenalg}
  Let $M$ be a $(*,\Delta)$-matrix of size $m \times n$. A forbidden
  submatrix from $\Tuckerset$ in $M$ that has a minimum number of rows
  can be found in $O(\Delta^3m^2n\dot(m+n^2))$ time. Within the same
  time, one can also find a forbidden submatrix from $\Tuckerset$ in
  $M$ that has a minimum number of columns, a minimum number of rows
  and columns, or a minimum number of entries.
\end{theoremsansproof}

The procedure used to prove Theorem~\ref{th:forbiddenalg} is as
follows. 
\begin{enumerate}
\item Run the algorithm mentioned in
  Propostition~\ref{prop:tuckerdomapprox} to obtain a submatrix $A$
  which is an approximation of the smallest forbidden submatrix. (If
  the actual solution is $M_{I_k}$ or $M_{II_k}$ this will be an exact
  solution.)
\item Run the algorithm from Propostition~\ref{cor:miiitime} to find
  an induced $M_{III_k}$ and let this be $B$.
\item Run the algorithm from Corollary~\ref{cor:mivmvpolytime} twice
  -- once each for finding $M_{IV}$ and finding $M_{V}$. Let the
  results be $C$ and $D$ respectively.
\item Return the minimum size matrix among $A, B, C, D$ (minimum size
  being one of the following: number of rows, number of columns, sum
  of number of rows and columns, number of {\un} entries).
\end{enumerate}

In the next section we see yet another forbidden substructure for
\COP which is easier to compute.

\subsection{Incompatibility graph - a certificate for no \COP}
\label{sec:incompatibilitygr}

When a \COP test algorithm reports negative, it is hard to verify
this. While \cite{at72} gives forbidden substructures for \COP, there
are no efficient algorithms to find these strutures in a matrix. In
fact, \cite{v82} shows that even finding a forbidden submatrix with at
least $k$ rows in an input matrix is NP-complete. This was proven by
polynomially reducing the longest path problem for graphs to this
problem.

\cite{mcc04} describes another characterization for matrices without
\COP which is efficiently computable. This is done by way of an {\em
  incompatibility graph} of a set system (since \COP in
  matrices and set systems are equivalent problems as seen in
  Section~\ref{sec:surveycertalgo} we will use the term set system in
  this section, in line with \cite{mcc04}'s convention). The
construction of this graph is based on the following observation.

\begin{observation} 
  Consider elements $a, b, c \in U$ and a set system $\cF$ with
  universe $U$. If there is at least one set $S \in \cF$ \stt $a, c
  \in S$ but $b \notin S$, then it is impossible to have a \COP order
  that will place $b$ in between $a$ and $c$. \tnote[important]{also cite the
    inbetweenness problem}
\end{observation}

For the sake of argument, suppose there exists such a \COP
order. Clearly, it will not succeed in mapping $S$ to an interval
since $b \notin S$ which is a contradiction to the assumption that
this is a \COP order.  The ordered pair $(x,y)$ denotes that in the
\COP order $y$ comes after $x$. Based on above observations, this
means $(a,b)$ and $(b,c)$ are not compatible in any \COP order. So are
$(c,b)$ and $(b,a)$.  Thus a binary relation {\em is incompatible
  with} on $U \times U \setminus \{(a,a) \mid a \in U\}$ can be
defined based on the membership of elements of $U$ in the sets in
$\cF$.  \cite{mcc04} creates an incompatiblity graph with vertices
denoting ordered pairs and edges denoting the incompatibility
relation.

\begin{definition}[{\em Incompatibility graph \cite[Def.~6.1]{mcc04}}]
  \label{def:incompatibility}
  Let $\cF$ be a set system with universe $U$. Let $A_{\cF} = \{(a,b)
  \mid a, b \in U, a \ne b\}$. The {\em incompatibility graph}
  $G_I^\cF$ of $\cF$ is an undirected graph defined as follows.
%  \singlespacing
  \begin{enumerate}
  \item   The vertex set of $G_I^\cF$ is $ A_{\cF}$
  \item  The edge set of $G_I^\cF$ are pairs of the following forms
    \begin{itemize}
    \item $\{(a,b),(b,a)\}$ for all $a, b \in U$
    \item $\{(a,b),(b,c)\}$ if there exists $S \in \cF$ \stt $a, c
      \in S$ and $b \notin S$, for all $a, b, c \in U$
    \end{itemize}
  \end{enumerate}
  \dstop
\end{definition}

Thus an edge between two vertices $(a,b), (b,c)$ in $G_I^{\cF}$ means
that there exists no linear order on $U$ that can place $b$ after $a$
and place $c$ after $b$.  For any \COP order, there must not be any
incompatible pairs. Thus it must consist of an independent set $I$ of
the incompatibility graph. Moreover the independent set must have
exactly half the vertex set of $G_I^\cF$ since the \COP order must
involve all elements in $U$. Secondly, the reverse of a \COP order is
also a \COP order. Thus $A_\cF \setminus I$ must also be an
independent set. In other words, $G_I^\cF$ must be bipartite with
partitions being of equal size if $\cF$ has \COP. An odd cycle is a
forbidden structure in a bipartite graph, hence the same is forbidden
in the incompatibility graph of a set system with \COP. In other
words, if the set system has no \COP, its incompatibility graph must
have an odd cycle.  There is also a requirement on the length of this
odd cycle which is formally stated in Theorem~\ref{th:mccbipartite}.

\begin{theoremsansproof}[{\cite[Th.~6.1]{mcc04}, \cite[Corrected by
    Th.~2.6, Proof p. 44--47]{d08phd}}] 
  \label{th:mccbipartite}
  Let $\cF$ be a set family with universe $U$. Then $\cF$ has \COP
  \iff its incompatibility graph is bipartite and if it does not have
  the \COP, the incompatibility graph has an odd cycle of length at
  most $n+3$.
\end{theoremsansproof}

Searching for an odd cycle to recognize a bipartite graph is a known
and polynomially solved problem. \cite{mcc04}'s algorithm also labels
each edge of the cycle with a set from $\cF$ that documents the
incompatibility.

% \tnote[dire]{how does it give clues
%   about the $R$ node? See dom.}

%     \cite{mcc04} describes a different approach to COT. While all
%     previous COT algorithms gave the COP order if the matrix has the
%     property but exited stating negative if otherwise, this algorithm
%     gives an evidence by way of a certificate of matrix even when it
%     has no COP. This enables a user to verify the algorithm's result
%     even when the answer is negative. This is significant from an
%     implementation perspective because automated program verification
%     is hard and manual verification is more viable. Hence having a
%     certificate reinforces an implementation's credibility. Note that
%     when the matrix {\em has} COP, the COP order is the certificate.
%     The internal machinery of this algorithm is related to the
%     weighted betweenness problem addressed\tnote{in what way??} in
%     \cite{co98}.  \tnote{expand on the COP order graph creation and it
%       having to be bipartite for M to have COP. and thus an odd cycle
%       being an evidence of no COP.}

% \tnote[dire]{ -- sect 4.1 in cite:d08phd has many results surveyed. hardness
%   results, approx. results. results are usually for a class of
%   matrices $(a,b)$ where number {\un}s in columns and rows are
%   restriced to $a$ and $b$ . -- problem of flipping at most $k$
%   entries of $M$ to make it attain COP. this is NP complete
%   cite:b75-phd%
% \\(1) scite:lb62 showed that interval graphs are chordal and
%   AT-free. %
% \\couple of lines refering to tucker's submatrices. refer
%   earlier section.

% Another problem is to find the minimum number of entries in the matrix
% that can be toggled to result in a matrix with COP.  \cite{v85}
% discusses approximation of {\sc COP Augmentation} which is the problem
% of changing of the minimum number of zero entries to {\un}s so that
% the resulting matrix has COP. As mentioned earlier, this problem is
% known to be NP complete due to \cite{b75-phd}.  \tnote[pressing]{where should this go?
%   cite|tz04 (approx submatrix with COP sparse matrices)}
% }

%%%%%%%%%%%%% listing package for pretty formatting BEGIN
% \lstloadlanguages{C}
% \lstset{language=C, commentstyle=\scriptsize,
%   numberstyle=\tiny, numbers=left,               % Line numbers
%   stepnumber=2,numbersep=5pt,firstnumber=10, 
%   showstringspaces=false,                    % No explicit spaces
%   emph={printf},emphstyle=\underbar,      % Specific formatting
%   emph=[2]{sum},emphstyle=[2]\color{blue},      % Sp format (multiple)
%   emph=[3]{i},emphstyle=[3]\color{red}      % Sp format (multiple)
%   }

%   \begin{singlespacing}
%     % Text inside "lstlisting" env is verbatim.
%     \begin{lstlisting}[keywordstyle=\textbf]      
% int sum;
% int i; /* for loop variable */ 
% sum = 0; 
% for (i=0, i<n; i++) { 
%   sum += a[i]; 
% } 
% printf{"This is stupid."}
%     \end{lstlisting}
%   \end{singlespacing}

%%%%%%%%%%%%% listing package for pretty formatting END