### 2012-02-28 src/002ack.tex
finished proof reading. added ack.

\xclearpage
\acknowledgements

Like any body of work, big or small, this thesis and the study behind
it could not have been done by me without the support of many people
and things professionally and personally. I believe everyone and
everything I encountered in the last two years and seven months has directly or
indirectly helped me conclude this work. Following are the most
obvious and direct help that I explicitly acknowledge.

I have deep gratitude for my research advisor
Dr.\;N.\;S.\;Narayanaswamy. I have been very fortunate to be
associated with a teacher with such remarkable patience and tenacity
with his students that brings out the possible best in them. He let us
work with the right kind of liberty that is needed to discover and
practice one's own method of research and discipline.  I would like to
thank him much for all his brilliant ideas, many discussions (at times
repeated and with no judgement) and perpetual words of
encouragement. I doubt if I would have learned my research without his
persevering guidance.  I would like to thank all fellow students in my
research group for frequent technical discussions and faithful
reviews; also my friends outside AIDB lab, for all caffeine-induced
discussions and de-stressing arguments.

Institute of Mathematical Sciences, for all the mind-bending lectures
that set a foundation to my research. I would especially like to thank
Prof.\;R.\;Ramanujam for motivating and encouraging my interest in
interdisciplinary research.

My husband, Zabil, for his amazing support and encouragement to take a
break from my career and pursue my higher education. Moreover, for
putting up patiently with an approximate solution to our two-body
problem in these years.

My parents, Srinivasan and Seena, for always being there and helping
me in ways only parents can help. My sister, Renu, for her precocious optimism. Devi and Rinchen, for being a
hotline for procrastination-induced panic attacks and reminding me
that I can indeed complete my M.\;S.\;! LMN -- Lakshmi, Mahathi and
Nari, for being the closest thing to a family on campus always and
especially whenever I defaulted to being a hermit (usually a famished
one).  Jai, J.\;K.\;, Shivani and Sourab for having an open door
policy for me any day any time that I was in Chennai and incentivizing
my visits with great fun and awesome food.

And last but certainly not the least, IITM and its wonderful campus
for inspiring me to unlearn my old methods of learning and helping me
discover and enjoy research.