### 2012-03-02 src/declarationtext.tex
additional files created for submission

%\certificate
\xclearpage% \cleardoublepage
\setcounter{page}{1}

%\def\certheadingstr{Thesis Certificate}%{\sc \centering Thesis Certificate}
\section*{Declaration by Research Scholar}

\vspace*{0.5in}

\noindent 
This is to declare that the thesis titled {\bf
  \mythesistitleNOTSYNPOSIS} submitted by me to the {\bf Indian
  Institute of Technology Madras}, for the award of the degree of {\bf
  \mydegree}, is a bona fide record of the research work done by me
under the supervision of {\bf Dr.~\myadvisor}, Associate Professor at
Department of \mydept, IIT Madras.  The contents of this thesis, in
full or in parts, have not been submitted and will not be submitted to
any other Institute or University for the award of any degree or
diploma.

\vspace*{1.5in}

\begin{singlespacing}
  \hspace*{-0.25in}
  \parbox{3in}{
    \noindent {Anju Srinivasan} \\
    \small
    \noindent Research Scholar} \hfill
  \parbox{1.25in}{               % {1.75in}
    \small
    \vspace*{0.75in}
    \begin{tabular}[h]{r}
    \noindent {Chennai}\\         
    \noindent {\mythesissubmissiondate} 
    \end{tabular}
  }
\end{singlespacing}

\vspace*{1.5in}

\begin{singlespacing}
  \hspace*{-0.25in}
  \parbox{3in}{
    \noindent{\em Forwarded:}\\
    \noindent {\bf Dr.~\myadvisor} \\
    \small
    \noindent Research Guide} 
\end{singlespacing}